---
author: "lea wölbert"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preparations

```{r libraries, message = FALSE}

library(rlist)
library(CellChat)
library(tidyr)

```

## Load Data

```{r load_convtables, message = FALSE}

hgnc_conv <- read.delim(file = "../../data/databases/hgnc_complete_set.txt", 
                        header = TRUE)

mgi_conv <- read.delim(
  file = "../../data//databases/HGNC_AllianceHomology.rpt.txt",
  header = TRUE)

```

```{r load_dbs, message = FALSE}

ccdb <- CellChatDB.mouse
mjdb <- read.csv(file = "../../data/databases/db_MJ.csv", 
                 header = TRUE, sep = ";")
nmdb <- read.csv(file = "../../data/databases/db_natmi.csv", 
                 header = TRUE, sep = ",")
scdb <- read.csv(file = "../../data/databases/db_scagecom_m.csv", 
                 header = TRUE, sep = ";")
rmdb <- read.csv(file = "../../data/databases/db_rnamagnet.csv", 
                 header = TRUE, sep = ";")

```

## Build the Scaffold

Use mjdb as basis for the scaffold.

```{r lrdb_scaffold, message = FALSE}

lrdb <- list(
  interactions = data.frame(interaction_pair = vector("character",
                                                      length = nrow(mjdb)),
    ligand_symbol = mjdb$Ligand_GeneSymbol,    
    ligand_ensembl_id = mjdb$Ligand_ENSEMBL,
    receptor_symbol = mjdb$Receptor_GeneSymbol,
    receptor_ensembl_id = mjdb$Receptor_ENSEMBL,
    interaction_type = vector("character", length = nrow(mjdb))
  ),
  sources = data.frame(interaction_pair = vector(mode = "character", 
                                                 length = nrow(mjdb)),
    CellChatDB.mouse_row = vector("character", length = nrow(mjdb)), 
    Mende.et.al_row = vector("character", length = nrow(mjdb)),
    RNA.Magnet_row = vector("character", length = nrow(mjdb)),
    interaction_type_source = vector("character", length = nrow(mjdb))
  ),
  ligands = data.frame(ligand_ensembl_id = mjdb$Ligand_ENSEMBL,
    ligand_symbol = mjdb$Ligand_GeneSymbol,
    ligand_hgnc_id = vector("character", length = nrow(mjdb)),
    ligand_hgnc_symbol_h = vector("character", length = nrow(mjdb))
  ),
  receptors = data.frame(receptor_ensembl_id = mjdb$Receptor_ENSEMBL,
    receptor_symbol = mjdb$Receptor_GeneSymbol,
    receptor_hgnc_id = vector("character", length = nrow(mjdb)),
    receptor_hgnc_symbol_h = vector("character", length = nrow(mjdb))
  )
)

```

Concatenate Ligands and Receptors to get interactions pairs.

```{r interaction_pairs, message = FALSE}

lrdb$interactions$interaction_pair <- paste(lrdb$interactions$ligand_symbol, 
                                            lrdb$interactions$receptor_symbol, 
                                            sep = "&")
lrdb$sources$interaction_pair <- paste(lrdb$interactions$ligand_symbol,
                                       lrdb$interactions$receptor_symbol, 
                                       sep = "&")
table(duplicated(lrdb$interactions$interaction_pair))

```

Add the source from mjdb.

```{r source_mjdb, message = FALSE}

# use a little trick by altering mjdb to also contain interaction pairs
mjdb$pair <- paste(mjdb$Ligand_GeneSymbol, mjdb$Receptor_GeneSymbol, sep = "&")

lrdb$sources$Mende.et.al_row <- paste(match(lrdb$interactions$interaction_pair, 
                                            mjdb$pair))

```

```{r checkpoint1_save, message = FALSE}

list.save(lrdb, file = "../../data/databases/lrdb_001.rds")

```

```{r checkpoint1_load, message = FALSE}

lrdb <- list.load(file = "../../data/databases/lrdb_001.rds")

```

# Expand using other DBs

## CellChat DB

### Prepare

First prepare the database for comparison.
For instance, the interaction name format must be changed and multimeric
interactions should be split into two/three monomeric interactions.

```{r prepare_ccdb, message = FALSE}

# prepare ccdb for comparison
ccdb <- CellChatDB.mouse

ccdb$interaction$original_row[1:2021] <- paste(1:2021)
ccdb$interaction <- ccdb$interaction[,c(3, 4, 10, 11, 12)]
nrow(ccdb$interaction)

# subset into two objects, one already only contains monomeric interactions,
ccdb_mono <- ccdb$interaction[-grep("[+]", 
                                    ccdb$interaction$interaction_name_2),]
nrow(ccdb_mono)
# the other contains multimeric interactions and will be split
ccdb_split<- ccdb$interaction[grep("[+]", 
                                   ccdb$interaction$interaction_name_2),]
nrow(ccdb_split)

```

For interactions that are already monomeric only the format must be changed.

```{r prepare_ccdb_mono, message = FALSE}

# six mono interactions with two ligands are filtered out for easier coding
ccdb_mono[grep("[[:alnum:]]+[_]+[[:alnum:]]+", ccdb_mono$ligand),]
ccdb_mono <- ccdb_mono[-grep("[[:alnum:]]+[_]+[[:alnum:]]+", ccdb_mono$ligand),]

# change interaction format to suit lrdb and add into new col
ccdb_mono$interaction_name <- gsub("  - ", "&", ccdb_mono$interaction_name_2 )

# change receptor name 
ccdb_mono$receptor <- gsub("[[:print:]]+[&]", "", ccdb_mono$interaction_name)

# remove unneccessary cols 
ccdb_mono <- ccdb_mono[,c(1,2,3,5,6)]

table(duplicated(ccdb_mono$interaction_name))
paste(nrow(ccdb_mono), "unique monomeric interactions")

```

Multimeric interactions must be adjusted in format and split.

```{r prepare_ccdb_split, message = FALSE}

# three interactions with multiple ligands are filtered out for easier coding 
ccdb_split[grep("[[:alnum:]]+[_]+[[:alnum:]]+", ccdb_split$ligand),]
ccdb_split <- ccdb_split[-grep("[[:alnum:]]+[_]+[[:alnum:]]+", 
                               ccdb_split$ligand),]

# change the interaction for splitting 
ccdb_split$interaction_name_2 <- gsub("[[:print:]]+[(]", "", 
                                      ccdb_split$interaction_name_2)
ccdb_split$interaction_name_2 <- gsub("[)]", "", 
                                      ccdb_split$interaction_name_2)
ccdb_split$interaction_name_2 <- gsub("[+]", "-", 
                                      ccdb_split$interaction_name_2)

# insert into receptor col and delete interaction_name_2 col
ccdb_split$receptor <- ccdb_split$interaction_name_2
ccdb_split <- ccdb_split[,c(1,2,3,5)]

# check that no ligand-multimers were used in ccdb_split
length(grep("_", ccdb_split$ligand))

# separate interactions according to receptor column only
ccdb_split <- separate_rows(ccdb_split, receptor,  sep = "-", convert = TRUE)
nrow(ccdb_split)

# add interaction pair in the same format as lrdb into new col
ccdb_split$interaction_name <- paste(ccdb_split$ligand, 
                                     ccdb_split$receptor, 
                                     sep = "&")

# check for duplicates
table(duplicated(ccdb_split$interaction_name))

```

There are a lot of duplicates because most receptor multimers contain the same 
receptor subunits multiple times. 
For instance, Tgfb1&Tgfbr1-Tgfbr2 and Tgfb1&Tgfbr1-Acvr1 will both result in 
one Tgfb1&Tgfbr1 pair, each.
These duplicates must be removed without losing information on the original 
row in the original database.

```{r prepare_ccdb_split_dups, message = FALSE}

# try to remove the duplicates without deleting the original rows

# treat as two different objects: duplicates and originals
ccdb_split_dups <- ccdb_split[duplicated(ccdb_split$interaction_name),]
ccdb_split_ori <- ccdb_split[duplicated(ccdb_split$interaction_name) == FALSE,]
# there are no duplications in the originals as intended
table(duplicated(ccdb_split_ori))
  
# check out the matching interactions
pos_ccdb_split_ori <- which(!is.na(match(ccdb_split_ori$interaction_name, 
                                         ccdb_split_dups$interaction_name)))
head(ccdb_split_ori[pos_ccdb_split_ori,])
# 407 of the originals are duplicated

pos_ccdb_split_dups <- which(!is.na(match(ccdb_split_dups$interaction_name, 
                                          ccdb_split_ori$interaction_name)))
head(ccdb_split_dups[pos_ccdb_split_dups,])
# of course all duplicates are found in the duplicated dataset
# this indicates some originals are duplicated multiple times
table(table(ccdb_split$interaction_name))

# these positions don't match perfectly so I can't just add everything
head(match(ccdb_split_ori$interaction_name[pos_ccdb_split_ori], 
           ccdb_split_dups$interaction_name[pos_ccdb_split_dups]))

# add them by loop 
for(i in ccdb_split_ori$interaction_name){
  match_pos <- match(ccdb_split_dups$interaction_name, i)
  match_pos_dup <- which(!is.na(match_pos))

  if(length(match_pos_dup) > 0){
    for(j in match_pos_dup){
      ccdb_split_ori$original_row[grep(
        i, ccdb_split_ori$interaction_name)] <- paste(
          ccdb_split_ori$original_row[grep(i, ccdb_split_ori$interaction_name)],
          ccdb_split_dups$original_row[j], sep = ", ")
     }
  }
}
  
ccdb_split <- ccdb_split_ori

paste(nrow(ccdb_split), "unique, monomeric interactions")

```

Combine the two database subsets again so comparison and addition of 
interactions must only be performed once.
Remove duplications were the monomeric interaction derived from a multimeric
interaction was already part of the database.

```{r combine_multi_mono, message = FALSE}

table(is.na(match(ccdb_mono$interaction_name, ccdb_split$interaction_name)))
# there are 23 interactions found in both split and mono
# most likely because a split multimeric interaction was already listed 
# as a monomeric interaction

# check out the matching interaction
pos_ccdb_mono <- which(!is.na(match(ccdb_mono$interaction_name, 
                                    ccdb_split$interaction_name)))
head(ccdb_mono[pos_ccdb_mono,])
pos_ccdb_split <- which(!is.na(match(ccdb_split$interaction_name, 
                                     ccdb_mono$interaction_name)))
head(ccdb_split[pos_ccdb_split,])
match(ccdb_mono$interaction_name[pos_ccdb_mono], 
      ccdb_split$interaction_name[pos_ccdb_split])
# they match perfectly

length(pos_ccdb_mono)
length(pos_ccdb_split)
# length is also the same since I got rid of duplicates in ccdb_split

# add the original_row from split to mono to enable tracing to both
# then delete these positions from split to avoid duplicates
ccdb_mono$original_row[pos_ccdb_mono] <- paste(
  ccdb_mono$original_row[pos_ccdb_mono], 
  ccdb_split$original_row[pos_ccdb_split], sep = ",")
head(ccdb_mono[pos_ccdb_mono,])                                         
ccdb_split <- ccdb_split[-pos_ccdb_split,]   

# combine both for comparison
ccdb_both <- rbind(ccdb_mono, ccdb_split )

# remove the duplicates which a
table(duplicated(ccdb_both$interaction_name))

paste(nrow(ccdb_both), "unique monomeric interactions")

```

### Compare and Add

Add all interactions which are not part of lrdb yet.

```{r compare_ccdb, message = FALSE}

table(is.na(match(ccdb_both$interaction_name, 
                  lrdb$interactions$interaction_pair)))
# around 1300 interactions form ccdb are not yet found in lrdb
table(is.na(match(ccdb_mono$interaction_name, 
                  lrdb$interactions$interaction_pair)))
# around 600 monomeric interactions form ccdb are not yet found in lrdb

```

Only add MONO by rbind but use info from BOTH for sources/interactions.


```{r add_ccdb_list, message = FALSE}

# vector of interaction pairs to be added
add_ccdb_pairs <- ccdb_mono$interaction_name[which(is.na(match(
  ccdb_mono$interaction_name, lrdb$interactions$interaction_pair)))]

# two variables for better visibility
ccdb_nr <- length(add_ccdb_pairs)
ccdb_pos <- which(is.na(match(ccdb_mono$interaction_name,
                              lrdb$interactions$interaction_pair)))

# don't add info on source, this will be added later for all shared interactions
add_ccdb <- list(
  interactions = data.frame(interaction_pair = add_ccdb_pairs,
    ligand_symbol = ccdb_mono$ligand[ccdb_pos],    
    ligand_ensembl_id = vector("character", length = ccdb_nr),
    receptor_symbol = ccdb_both$receptor[ccdb_pos],
    receptor_ensembl_id = vector("character", length = ccdb_nr),
    interaction_type = vector("character", length = ccdb_nr)
  ),
  sources = data.frame(interaction_pair = add_ccdb_pairs,
    CellChatDB.mouse_row = vector("character", length = ccdb_nr), 
    Mende.et.al_row = rep("NA", length = ccdb_nr),
    RNA.Magnet_row = vector("character", length = ccdb_nr),
    interaction_type_source = vector("character", length = ccdb_nr)
  ),
  ligands = data.frame(ligand_ensembl_id = vector("character", 
                                                  length = ccdb_nr),
    ligand_symbol = ccdb_mono$ligand[ccdb_pos],
    ligand_hgnc_id = vector("character", length = ccdb_nr),
    ligand_hgnc_symbol_h = vector("character", length = ccdb_nr)
  ),
  receptors = data.frame(receptor_ensembl_id = vector("character", 
                                                      length = ccdb_nr),
    receptor_symbol = ccdb_mono$receptor[ccdb_pos],
    receptor_hgnc_id = vector("character", length = ccdb_nr),
    receptor_hgnc_symbol_h = vector("character", length = ccdb_nr)
  )
)

head(add_ccdb)

```

```{r add_ccdb_rbind, message = FALSE}

lrdb$interactions <- rbind(lrdb$interactions, add_ccdb$interactions)
lrdb$sources <- rbind(lrdb$sources, add_ccdb$sources)
lrdb$ligands <- rbind(lrdb$ligands, add_ccdb$ligands)
lrdb$receptors <- rbind(lrdb$receptors, add_ccdb$receptors)

head(lrdb)
tail(lrdb)

table(is.na(match(ccdb_both$interaction_name, 
                  lrdb$interactions$interaction_pair)))
# now all pairs from ccdb_both are contained in lrdb

```

```{r add_ccdb_sourceinfo, message = FALSE}

# add info by matching between ccdb_both and lrdb

shared_pairs_ccdb <- ccdb_both$interaction_name[which(!is.na(match(
  ccdb_both$interaction_name, lrdb$interactions$interaction_pair)))]
length(shared_pairs_ccdb)

lrdb_pos <- which(!is.na(match(lrdb$interactions$interaction_pair, 
                               shared_pairs_ccdb)))
ccdb_both_pos <- which(!is.na(match(ccdb_both$interaction_name,
                               shared_pairs_ccdb)))
fillin_lrdb <- match(ccdb_both$interaction_name[ccdb_both_pos],
                     lrdb$interactions$interaction_pair[lrdb_pos])

# to check if they correspond
head(match(ccdb_both$interaction_name[ccdb_both_pos], 
      lrdb$interactions$interaction_pair[lrdb_pos][fillin_lrdb]))
# they match perfectly, so I can just fill in

# add to lrdb
lrdb$interactions$interaction_type[lrdb_pos][
  fillin_lrdb] <- ccdb_both$annotation[ccdb_both_pos]
lrdb$sources$CellChatDB.mouse_row[lrdb_pos][
  fillin_lrdb] <- ccdb_both$original_row[ccdb_both_pos]
lrdb$sources$interaction_type_source[lrdb_pos][
  fillin_lrdb] <- paste(lrdb$sources$interaction_type_source[lrdb_pos][
    fillin_lrdb], "CellChatDB.mouse")

table(duplicated(lrdb$interactions$interaction_pair))
tail(lrdb$interactions)
tail(lrdb$sources)

 
```

### Checkpoint

```{r checkpoint2_save, message = FALSE}

list.save(lrdb, file = "../../data/databases/lrdb_002.rds")

```

```{r checkpoint2_load, message = FALSE}

lrdb <- list.load(file = "../../data/databases/lrdb_002.rds")

```

## RNA Magnet

### Prepare

```{r prepare_rmdb, message = FALSE}

# prepare rmdb for comparison
rmdb <- read.csv(file = "../../data/databases/db_rnamagnet.csv", 
                 header = TRUE, sep = ";")
tail(rmdb)

rmdb$original_row[1:nrow(rmdb)] <- paste(1:nrow(rmdb))
tail(rmdb)
rmdb <- rmdb[,c(1, 2, 3, 5, 6, 9)]

# there are some manual annotations, including "Incorrect" and "Scavenger"
# I will not remove them here because I need them for reference

# subset into two objects, one already only contains monomeric interactions,
rmdb_mono <- rmdb[-grep("&", rmdb$Pair.Name),]
nrow(rmdb_mono)
# the other contains multimeric interactions and will be split
rmdb_split <- rmdb[grep("&", rmdb$Pair.Name),]
nrow(rmdb_split)

```

```{r prepare_rmdb_mono, message = FALSE}

# some interactions concatenated with |, remove these for easier coding
rmdb_mono <- rmdb_mono[-grep("[|]", rmdb_mono$Pair.Name),]

# change interaction format to suit lrdb and add into new col
rmdb_mono$Pair.Name <- gsub("-", "&", rmdb_mono$Pair.Name)

table(duplicated(rmdb_mono$Pair.Name))
paste(nrow(rmdb_mono), "unique monomeric interactions")

```

```{r prepare_rmdb_split, message = FALSE}

# some interactions concatenated with |, remove these for easier coding
rmdb_split <- rmdb_split[-grep("[|]", rmdb_split$Pair.Name),]
nrow(rmdb_split)

# one interaction involves more than two R or Ls, remove it for consistency
rmdb_split <- rmdb_split[-grep("[[:alnum:]]+[&]+[[:alnum:]]+[&]+[[:alnum:]]+",
                               rmdb_split$Pair.Name),]
nrow(rmdb_split)

# separate interactions according to receptor and Ligand column
rmdb_split <- separate_rows(rmdb_split,
                            Receptor.Mouse, 
                            Ligand.Mouse, 
                            sep = "&", 
                            convert = TRUE)
rmdb_split$Pair.Name2 <- paste(rmdb_split$Ligand.Mouse, 
                               rmdb_split$Receptor.Mouse, 
                               sep = "&")

# check for duplicates
table(duplicated(rmdb_split$Pair.Name2))
# the number of duplicates is low but must be attended to
table(table(rmdb_split$Pair.Name2))
paste("nrow after splitting:", nrow(rmdb_split))

```

```{r prepare_ccdb_split_dups, message = FALSE}

# try to remove the duplicates without deleting the original rows

# treat as two different objects: duplicates and originals
rmdb_split_dups <- rmdb_split[duplicated(rmdb_split$Pair.Name2),]
nrow(rmdb_split_dups)
rmdb_split_ori <- rmdb_split[duplicated(rmdb_split$Pair.Name2) == FALSE,]
nrow(rmdb_split_ori)

# there are no duplications in the originals as intended
table(duplicated(rmdb_split_ori))
  
# check out the matching interactions
pos_rmdb_split_ori <- which(!is.na(match(rmdb_split_ori$Pair.Name2, 
                                         rmdb_split_dups$Pair.Name2)))
length(pos_rmdb_split_ori)
head(rmdb_split_ori[pos_rmdb_split_ori,])
# 32 of the originals are duplicated

pos_rmdb_split_dups <- which(!is.na(match(rmdb_split_dups$Pair.Name2, 
                                          rmdb_split_ori$Pair.Name2)))
nrow(rmdb_split_dups)
length(pos_rmdb_split_dups)
head(rmdb_split_dups[pos_rmdb_split_dups,])
# of course all duplicates are found in the duplicated dataset
# this indicates some originals are duplicated multiple times

# these positions don't match perfectly so I can't just add everything
head(match(rmdb_split_ori$Pair.Name2[pos_rmdb_split_ori], 
           rmdb_split_dups$Pair.Name2[pos_rmdb_split_dups]))

# add them by loop 
for(i in rmdb_split_ori$Pair.Name2){
  match_pos <- match(rmdb_split_dups$Pair.Name2, i)
  match_pos_dup <- which(!is.na(match_pos))

  if(length(match_pos_dup) > 0){
    for(j in match_pos_dup){
      rmdb_split_ori$original_row[grep(
        i, rmdb_split_ori$Pair.Name2)] <- paste(
          rmdb_split_ori$original_row[grep(i, rmdb_split_ori$Pair.Name2)],
          rmdb_split_dups$original_row[j], sep = ", ")
     }
  }
}
  
rmdb_split <- rmdb_split_ori
table(duplicated(rmdb_split$Pair.Name2))

paste(nrow(rmdb_split), "unique, monomeric interactions")

```

```{r combine_multi_mono, message = FALSE}

# adjust rmdb_mono to rmdb_split
rmdb_mono$Pair.Name2 <- rmdb_mono$Pair.Name

table(is.na(match(rmdb_mono$Pair.Name2, rmdb_split$Pair.Name2)))
# there are 5 interactions found in both split and mono
# most likely because a split multimeric interaction was already listed 
# as a monomeric interaction

# check out the matching interaction
pos_rmdb_mono <- which(!is.na(match(rmdb_mono$Pair.Name2, 
                                    rmdb_split$Pair.Name2)))
rmdb_mono[pos_rmdb_mono,]
pos_rmdb_split <- which(!is.na(match(rmdb_split$Pair.Name2, 
                                     rmdb_mono$Pair.Name2)))
rmdb_split[pos_rmdb_split,]
rmdb_split_fill_pos <- match(rmdb_mono$Pair.Name2[pos_rmdb_mono], 
      rmdb_split$Pair.Name2[pos_rmdb_split])
# they match perfectly

length(pos_rmdb_mono)
length(pos_rmdb_split)
# length is also the same since I got rid of duplicates in ccdb_split

# add the original_row from split to mono to enable tracing to both
# then delete these positions from split to avoid duplicates
rmdb_mono$original_row[pos_rmdb_mono][rmdb_split_fill_pos] <- paste(
  rmdb_mono$original_row[pos_rmdb_mono], 
  rmdb_split$original_row[pos_rmdb_split], sep = ",")
rmdb_mono[pos_rmdb_mono,]                                       
rmdb_split <- rmdb_split[-pos_rmdb_split,]   

# combine both for comparison
rmdb_both <- rbind(rmdb_mono, rmdb_split)

# remove the duplicates which a
table(duplicated(rmdb_both$Pair.Name2))

paste(nrow(rmdb_both), "unique monomeric interactions")

```

### Compare and Add

Add all interactions which are not part of lrdb yet.

```{r compare_rmdb, message = FALSE}

table(is.na(match(rmdb_both$Pair.Name2, 
                  lrdb$interactions$interaction_pair)))
# around 650 interactions form rmdb are not yet found in lrdb
# this includes "incorrect" or "scavenger" interactions

```

```{r add_ccdb_list, message = FALSE}

# vector of interaction pairs to be added
add_rmdb_pairs <- rmdb_both$Pair.Name2[which(is.na(match(
  rmdb_both$Pair.Name2, lrdb$interactions$interaction_pair)))]

# two variables for better visibility
rmdb_nr <- length(add_rmdb_pairs)
rmdb_pos <- which(is.na(match(rmdb_both$Pair.Name2,
                              lrdb$interactions$interaction_pair)))

# don't add info on source, this will be added later for all shared interactions
add_rmdb <- list(
  interactions = data.frame(interaction_pair = add_rmdb_pairs,
    ligand_symbol = rmdb_both$Ligand.Mouse[rmdb_pos],    
    ligand_ensembl_id = vector("character", length = rmdb_nr),
    receptor_symbol = rmdb_both$Receptor.Mouse[rmdb_pos],
    receptor_ensembl_id = vector("character", length = rmdb_nr),
    interaction_type = vector("character", length = rmdb_nr)
  ),
  sources = data.frame(interaction_pair = add_rmdb_pairs,
    CellChatDB.mouse_row = rep("NA", length = rmdb_nr), 
    Mende.et.al_row = rep("NA", length = rmdb_nr),
    RNA.Magnet_row = vector("character", length = rmdb_nr),
    interaction_type_source = vector("character", length = rmdb_nr)
  ),
  ligands = data.frame(ligand_ensembl_id = vector("character", 
                                                  length = rmdb_nr),
    ligand_symbol = rmdb_both$Ligand.Mouse[rmdb_pos],
    ligand_hgnc_id = vector("character", length = rmdb_nr),
    ligand_hgnc_symbol_h = vector("character", length = rmdb_nr)
  ),
  receptors = data.frame(receptor_ensembl_id = vector("character", 
                                                      length = rmdb_nr),
    receptor_symbol = rmdb_both$Receptor.Mouse[rmdb_pos],
    receptor_hgnc_id = vector("character", length = rmdb_nr),
    receptor_hgnc_symbol_h = vector("character", length = rmdb_nr)
  ),
  annotation = data.frame(
    interaction_pair = add_rmdb_pairs,
    Manual.Annotation =  vector("character", length = rmdb_nr)
  )
)

# add manial annotation from lrdb, it will later be removed though

tail(add_rmdb$interactions)
table(duplicated(add_rmdb$interactions$interaction_pair))

```

```{r add_rmdb_rbind, message = FALSE}

# add annotations in an extra df, it will be deleted later
nrow(lrdb$interactions)
lrdb$annotation <- data.frame(
  interaction_pair = lrdb$interactions$interaction_pair,
  Manual.Annotation = vector(mode = "character", 
                             length = nrow(lrdb$interactions)))
lrdb$interactions <- rbind(lrdb$interactions, add_rmdb$interactions)
lrdb$sources <- rbind(lrdb$sources, add_rmdb$sources)
lrdb$ligands <- rbind(lrdb$ligands, add_rmdb$ligands)
lrdb$receptors <- rbind(lrdb$receptors, add_rmdb$receptors)
lrdb$annotation <- rbind(lrdb$annotation, add_rmdb$annotation)
nrow(lrdb$interactions)

tail(lrdb$interactions)
tail(lrdb$sources)
tail(lrdb$annotation)

table(duplicated(lrdb$interactions$interaction_pair))
table(is.na(match(rmdb_both$Pair.Name2, 
                  lrdb$interactions$interaction_pair)))
# now all pairs from ccdb_both are contained in lrdb

```

```{r add_rmdb_sourceinfo, message = FALSE}

# add info by matching between ccdb_both and lrdb
shared_pairs_rmdb <- rmdb_both$Pair.Name2[which(!is.na(match(
  rmdb_both$Pair.Name2, lrdb$interactions$interaction_pair)))]
length(shared_pairs_rmdb)

lrdb_pos2 <- which(!is.na(match(lrdb$interactions$interaction_pair, 
                               shared_pairs_rmdb)))
rmdb_both_pos <- which(!is.na(match(rmdb_both$Pair.Name2,
                               shared_pairs_rmdb)))
fillin_lrdb <- match(rmdb_both$Pair.Name2[rmdb_both_pos],
                     lrdb$interactions$interaction_pair[lrdb_pos2])

# to check if they correspond
head(match(rmdb_both$Pair.Name2[rmdb_both_pos], 
      lrdb$interactions$interaction_pair[lrdb_pos2][fillin_lrdb]))
tail(match(rmdb_both$Pair.Name2[rmdb_both_pos], 
      lrdb$interactions$interaction_pair[lrdb_pos2][fillin_lrdb]))
# they match perfectly, so I can just fill in

# add to lrdb
lrdb$interactions$interaction_type[lrdb_pos2][
  fillin_lrdb] <- paste(lrdb$interactions$interaction_type[lrdb_pos2][
  fillin_lrdb], rmdb_both$Ligand.CC[rmdb_both_pos], sep = ",")
lrdb$sources$RNA.Magnet_row[lrdb_pos2][
  fillin_lrdb] <- rmdb_both$original_row[rmdb_both_pos]
lrdb$sources$interaction_type_source[lrdb_pos2][
  fillin_lrdb] <- paste(lrdb$sources$interaction_type_source[lrdb_pos2][
    fillin_lrdb], "RNA.Magnet", sep = ";")
lrdb$annotation$Manual.Annotation[lrdb_pos2][
  fillin_lrdb] <- rmdb_both$ManualAnnotation[rmdb_both_pos]

table(duplicated(lrdb$interactions$interaction_pair))
head(lrdb$interactions)
head(lrdb$sources)
tail(lrdb$interactions)
tail(lrdb$sources)
tail(lrdb$annotation)
unique(lrdb$annotation$Manual.Annotation)

```

### Check Incorrect/Scavenger

There are manual annotations in rmdb to help evaluate the usefulness/
correctnes of a certain interaction.

- Correct: the exact interactions they used for their paper

- Incorrect: no evidence of interaction

- Irrelevant: interactions involving activated complement system components

- Scavenger: scavenger receptors

- Not Expressed: not expressed in Bone Marrow

- Likely: likely expressed?

So far, all interactions and annotations in lrdb were kept for reference. 
Now, all "incorrect" and "scavenger" interactions are removed.

```{r remove_inc_rmdb, message = FALSE}

remove_pos <- c(grep("Incorrect", lrdb$annotation$Manual.Annotation),
                grep("Scavenger", lrdb$annotation$Manual.Annotation))

# check
lrdb$annotation[remove_pos,]
lrdb$sources[remove_pos,]

# remove
lrdb$interactions <- lrdb$interactions[-remove_pos,]
lrdb$sources <- lrdb$sources[-remove_pos,]
lrdb$ligands <- lrdb$ligands[-remove_pos,]
lrdb$receptors <- lrdb$receptors[-remove_pos,]

# make a "new" lrdb list not containing annotations
lrdb <- list(interactions = lrdb$interactions,
             sources = lrdb$sources,
             ligands = lrdb$ligands,
             receptors = lrdb$receptors)

table(duplicated(lrdb$interactions$interaction_pair))

# also add NAs instead of "" into sources
lrdb$sources$RNA.Magnet_row[
  lrdb$sources$RNA.Magnet_row == ""] <- paste("NA")
lrdb$sources$CellChatDB.mouse_row[
  lrdb$sources$CellChatDB.mouse_row == ""] <- paste("NA")

```

### Checkpoint

```{r checkpoint3_save, message = FALSE}

list.save(lrdb, file = "../../data/databases/lrdb_003.rds")

```

```{r checkpoint3_load, message = FALSE}

lrdb <- list.load(file = "../../data/databases/lrdb_003.rds")

```

### Check some info/metrics

```{r checkpoint3_check_itype, message = FALSE}

# Annotations of interacton type

# how many interactions have missing annotations?
table(is.na(lrdb$interactions$interaction_type))
table(lrdb$interaction$interaction_type == "")
# no nas but around 400 of 3000 interactions have no annotations

# overlap of interaction type annotations
table(lrdb$interaction$interaction_type[
  lrdb$interaction$interaction_type != ""])

# get the positions of conflicting annotations
conflict_pos <- c(
  grep("Cell-Cell Contact,Secreted", lrdb$interaction$interaction_type),
  grep("ECM-Receptor,Secreted", lrdb$interaction$interaction_type),
  grep("Secreted Signaling,Membrane", lrdb$interaction$interaction_type),
  grep("Secreted Signaling,ECM", lrdb$interaction$interaction_type)
 )

# only 37 conflicting annotations
# these can be manually checked later
length(conflict_pos)

lrdb$interactions[conflict_pos,]
lrdb$sources[conflict_pos,]

```

Further annotations may be added manually or from other sources.
Depending on how many non-annotated interactions are even found when compared
to insp_ref_shared.

```{r checkpoint3_check_lris, message = FALSE}

# check that all lrdb interaction pairs are identical

table(is.na(match(lrdb$interactions$interaction_pair, 
                  lrdb$sources$interaction_pair)))
table(is.na(match(lrdb$sources$interaction_pair, 
                  lrdb$interactions$interaction_pair)))

# yes, they are

```


```{r  checkpoint3_check_doublei, message = FALSE}

# check that no double interactions are existent (except for ECM)

# make a vector of "turned around" interactions

ta_ints <- paste(lrdb$interactions$receptor_symbol,
                 lrdb$interactions$ligand_symbol,
                 sep = "&")

table(is.na(match(ta_ints, lrdb$interactions$interaction_pair)))
# around 570 interactions also appear turned around
# are these by chance ECM? (that would be ok)

lrdb$interactions[which(!is.na(match(lrdb$interactions$interaction_pair,
                                     ta_ints))),]
unique(lrdb$interactions$interaction_type[which(!is.na(match(
  lrdb$interactions$interaction_pair, ta_ints)))])

# weird cases (these and their counterparts)
# own knowledge/doubts?
grep("Sfrp1&Wnt3a", lrdb$interactions$interaction_pair)

# conflicting annotations
grep("Cd27&Cd70", lrdb$interactions$interaction_pair)
grep("Fpr2&Anxa1", lrdb$interactions$interaction_pair)
grep("Fpr1&Anxa1", lrdb$interactions$interaction_pair)
# guess TNF receptors have a similar problem as ErbbRs
grep("Tnfrsf10b&Tnfsf10", lrdb$interactions$interaction_pair)
grep("Tnfrsf4&Tnfsf4", lrdb$interactions$interaction_pair)
grep("Tnfrsf9&Tnfsf9", lrdb$interactions$interaction_pair)
grep("Tnfrsf8&Tnfsf8", lrdb$interactions$interaction_pair)
grep("Tnfrsf25&Tnfsf15", lrdb$interactions$interaction_pair)
grep("Tnfrsf14&Btla", lrdb$interactions$interaction_pair)
grep("Tnfrsf18&Tnfsf18", lrdb$interactions$interaction_pair)
# Egfr cases are a little bit difficult, probably caused by
# annotation of dimerization of ErbRs and inappropriate splitting
grep("Egfr&Egf", lrdb$interactions$interaction_pair)
grep("Erbb4&Egf", lrdb$interactions$interaction_pair)
grep("Erbb2&Egf", lrdb$interactions$interaction_pair)
grep("Erbb3&Egf", lrdb$interactions$interaction_pair)
grep("Erbb3&Areg", lrdb$interactions$interaction_pair)
grep("Egfr&Areg", lrdb$interactions$interaction_pair)
grep("Egfr&Epgn", lrdb$interactions$interaction_pair)

# receptor = ligand
grep("Ptprs&Ptprs", lrdb$interactions$interaction_pair)
grep("Il17a&Il17a", lrdb$interactions$interaction_pair)
grep("Fga&Fga", lrdb$interactions$interaction_pair)
grep("Sfrp1&Sfrp1", lrdb$interactions$interaction_pair)
grep("Thy1&Thy1", lrdb$interactions$interaction_pair)

# "secretion" annotation
grep("Cdh5&Fga", lrdb$interactions$interaction_pair)

```

These few interactions (and their counterparts) will be checked manually 
for their correctness if found in other databases or insp_ref_shared. 
Interactions that are not found anywhere else will be discarded.

```{r checkpoint3_check_symbols, message = FALSE}

# check if the symbols of the imported interaction partners that do not
# have an ensembl ID corresponds to the official symbols
table(is.na(match(lrdb$ligands$ligand_symbol, mgi_conv$MGI.Accession.ID)))
table(is.na(match(lrdb$receptors$receptor_symbol, mgi_conv$MGI.Accession.ID)))
# a low number of "symbols in lrdb does not correspond to actual mgi symbols
# these will most likely be discarded because they cannot be mapped 
# unambiguously unless some cases are really important 

# LIGANDS:
# exclude ligands that already have an ID from mjdb
pos_lig <- lrdb$ligands$ligand_ensembl_id[which(is.na(match(
  lrdb$ligands$ligand_symbol, mgi_conv$MGI.Accession.ID)))] == ""

lrdb$ligands[which(is.na(match(lrdb$ligands$ligand_symbol, 
                               mgi_conv$MGI.Accession.ID))),][pos_lig,]
lrdb$sources[which(is.na(match(lrdb$ligands$ligand_symbol, 
                               mgi_conv$MGI.Accession.ID))),][pos_lig ,]
unique(lrdb$ligands$ligand_symbol[which(is.na(match(lrdb$ligands$ligand_symbol, 
                               mgi_conv$MGI.Accession.ID)))][pos_lig])
# around 30 ligands in total cannot be found in mgi_conv

# RECEPTORS:
pos_rec <- lrdb$receptors$receptor_ensembl_id[which(is.na(match(
  lrdb$receptors$receptor_symbol, mgi_conv$MGI.Accession.ID)))] == ""

lrdb$receptors[which(is.na(match(lrdb$receptors$receptor_symbol, 
                               mgi_conv$MGI.Accession.ID))),][pos_rec,]
lrdb$sources[which(is.na(match(lrdb$receptors$receptor_symbol, 
                               mgi_conv$MGI.Accession.ID))),][pos_rec ,]
unique(lrdb$receptors$receptor_symbol[which(is.na(match(
  lrdb$receptors$receptor_symbol, mgi_conv$MGI.Accession.ID)))][pos_rec])
# around 20 receptors in total cannot be found in mgi_conv

# check if some of these Ls or Rs are important and need some manual annotation

```
