---
author: "lea wölbert"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Preparations

```{r libraries, message = FALSE}

library(rlist)
library(CellChat)
library(tidyr)

```

## Load Data

```{r load_convtables, message = FALSE}

hgnc_conv <- read.delim(file = "../../data/databases/hgnc_complete_set.txt", 
                        header = TRUE)

mgi_conv <- read.delim(
  file = "../../data//databases/HGNC_AllianceHomology.rpt.txt",
  header = TRUE)

```

```{r load_dbs, message = FALSE}

ccdb <- CellChatDB.mouse
mjdb <- read.csv(file = "../../data/databases/db_MJ.csv", 
                 header = TRUE, sep = ";")
nmdb <- read.csv(file = "../../data/databases/db_natmi.csv", 
                 header = TRUE, sep = ",")
scdb <- read.csv(file = "../../data/databases/db_scagecom_m.csv", 
                 header = TRUE, sep = ";")
rmdb <- read.csv(file = "../../data/databases/db_rnamagnet.csv", 
                 header = TRUE, sep = ";")

```

```{r load_lrdb, message = FALSE}

lrdb <- list.load(file = "../../data/databases/lrdb_003.rds")

```

```{r load_lrdb, message = FALSE}

insp_ref <- list.load(file = "../../data/databases/insp_ref_comp.rds")
insp_ref_shared <- list.load(file = "../../data/databases/insp_ref_shared.rds")

```


# Compare to remaining DBs
 
To get an idea if the interactions contain more or less what is required,
compare to the two remaining databases that were not used as source.

I will make addition of further data dependent on mapping to insp_ref_shared,
since only genes contained in that database are of interest further on.

## SCDB

```{r compare_scdb, message = FALSE}

# change scdb to make comparison easier
scdb_comp <- scdb
scdb_comp$LRI <- gsub(":", "&", scdb_comp$LRI)

```

```{r compare_scdb_mono, message = FALSE}

# create df  with only monomeric interactions from scdb
scdb_comp_mono <- scdb_comp[scdb_comp$LIGAND_2 == "",]
scdb_comp_mono <- scdb_comp_mono[scdb_comp_mono$RECEPTOR_2 == "",]
scdb_comp_mono <- scdb_comp_mono[scdb_comp_mono$RECEPTOR_2 == "",]

# compare with lrdb
table(is.na(match(lrdb$interactions$interaction_pair, scdb_comp_mono$LRI)))
table(is.na(match(scdb_comp_mono$LRI, lrdb$interactions$interaction_pair)))
# around 1800 monomeric interactions are shared
# both dbs have around 1800 monomeric interactions not found in the other

# subset into non-matched interactions
scdb_comp_mono_na <- scdb_comp_mono
scdb_comp_mono_na <- scdb_comp_mono_na[which(is.na(match(
  scdb_comp_mono$LRI, lrdb$interactions$interaction_pair))),]
scdb_comp_mono_na <- scdb_comp_mono_na[, c(1,2,4)]
  
# compare to insp_ref_shared
scdb_comp_mono_na$ligand_id <- mgi_conv$NCBI.Gene.strand[
  match(scdb_comp_mono_na$LIGAND_1, mgi_conv$MGI.Accession.ID)]

scdb_comp_mono_na$receptor_id <- mgi_conv$NCBI.Gene.strand[
  match(scdb_comp_mono_na$RECEPTOR_1, mgi_conv$MGI.Accession.ID)]

table(is.na(match(scdb_comp_mono_na$ligand_id, 
                  insp_ref_shared$ids$ensembl_mbl6_id)))
table(is.na(match(scdb_comp_mono_na$receptor_id, 
                  insp_ref_shared$ids$ensembl_mbl6_id)))
# around 1400 ligand and receptor IDs are found in insp_ref_shared
pos_lig_shared <- which(!is.na(match(scdb_comp_mono_na$ligand_id, 
                  insp_ref_shared$ids$ensembl_mbl6_id)))
pos_rec_shared <- which(!is.na(match(scdb_comp_mono_na$receptor_id, 
                  insp_ref_shared$ids$ensembl_mbl6_id)))
table(is.na(match(pos_lig_shared, pos_rec_shared)))
table(is.na(match(pos_rec_shared, pos_lig_shared)))
# a total of around 1170 interactions has ids in insp_ref_shared for
# both ligand and receptors and is not found in lrdb

# subset into insp_ref_shared-contained monomeric non-matching interactions
shared_scdb_comp_mono_na <- scdb_comp_mono_na
shared_scdb_comp_mono_na <- shared_scdb_comp_mono_na[
  which(!is.na(match(pos_lig_shared, pos_rec_shared))),]
table(duplicated(shared_scdb_comp_mono_na$LRI))

# could add around 1500 new unique monomeric interactions
# think about whether I should add those since they're not annotated at all
# and may not be significant to BM

```


```{r compare_scdb_multi_dfs, message = FALSE}
# create df  with only monomeric interactions from scdb
# make different dfs for different situations
multi_pos_llr <- c(which(scdb_comp$LIGAND_2 != ""))

multi_pos_lrr <- c(which(scdb_comp$LIGAND_2 != ""),
                   which(scdb_comp$RECEPTOR_2 == ""),
                   which(scdb_comp$RECEPTOR_3 != ""))
multi_pos_lrrr <- c(which(scdb_comp$LIGAND_2 != ""),
                    which(scdb_comp$RECEPTOR_2 == ""),
                    which(scdb_comp$RECEPTOR_3 == ""))
multi_pos_llrr <- c(which(scdb_comp$LIGAND_2 == ""),
                    which(scdb_comp$RECEPTOR_2 == ""),
                    which(scdb_comp$RECEPTOR_3 != ""))
multi_pos_llrrr <- c(which(scdb_comp$LIGAND_2 != ""),
                     which(scdb_comp$RECEPTOR_2 != ""),
                     which(scdb_comp$RECEPTOR_3 != ""))                   

scdb_multi_llr <- scdb_comp[multi_pos_llr,]
scdb_multi_llrrr <- scdb_multi_llr[scdb_multi_llr$RECEPTOR_3 != "",]
scdb_multi_llrr <- scdb_multi_llr[scdb_multi_llr$RECEPTOR_2 != "",]
scdb_multi_llrr <- scdb_multi_llrr[scdb_multi_llrr$RECEPTOR_3 == "",]
scdb_multi_llr <- scdb_multi_llr[scdb_multi_llr$RECEPTOR_2 == "",]

nrow(scdb_multi_llr)
nrow(scdb_multi_llrr)
nrow(scdb_multi_llrrr)

scdb_multi_lrr <- scdb_comp[-multi_pos_llr,]
scdb_multi_lrr <- scdb_multi_lrr[scdb_multi_lrr$RECEPTOR_2 != "",]
scdb_multi_lrr <- scdb_multi_lrr[scdb_multi_lrr$RECEPTOR_3 == "",]
scdb_multi_lrrr <- scdb_multi_lrr[scdb_multi_lrr$RECEPTOR_3 != "",]

nrow(scdb_multi_lrr)
nrow(scdb_multi_lrrr)

```

The LRR interactions are by far the most numerous.

```{r split_llr, message = FALSE}

# split llr
scdb_multi_llr_split <- separate_rows(scdb_multi_llr, LRI, sep = "_")
scdb_multi_llr_split$LRI[-grep("&", scdb_multi_llr_split$LRI)] <- paste(
  scdb_multi_llr_split$LRI[-grep("&", scdb_multi_llr_split$LRI)], 
  scdb_multi_llr_split$RECEPTOR_1[-grep("&", scdb_multi_llr_split$LRI)], 
  sep = "&")

table(duplicated(scdb_multi_llr_split$LRI))
table(is.na(match(scdb_multi_llr_split$LRI, 
                  lrdb$interactions$interaction_pair)))
table(is.na(match(lrdb$interactions$interaction_pair, 
                  scdb_multi_llr_split$LRI)))
# 8 of those interactions are not in lrdb
# so that doesn't give much benefit

```

```{r split_llrr, message = FALSE}

# split llrr
scdb_multi_llrr_split <- scdb_multi_llrr
scdb_multi_llrr_split$lri1 <- paste(scdb_multi_llrr$LIGAND_1, 
                              scdb_multi_llrr$RECEPTOR_1, sep = "&")
scdb_multi_llrr_split$lri2 <- paste(scdb_multi_llrr$LIGAND_1, 
                              scdb_multi_llrr$RECEPTOR_2, sep = "&")
scdb_multi_llrr_split$lri3 <- paste(scdb_multi_llrr$LIGAND_2, 
                              scdb_multi_llrr$RECEPTOR_1, sep = "&")
scdb_multi_llrr_split$lri4 <- paste(scdb_multi_llrr$LIGAND_2, 
                              scdb_multi_llrr$RECEPTOR_2, sep = "&")

scdb_multi_llrr_split$lri1[duplicated(scdb_multi_llrr_split$lri1) == TRUE] <- ""
scdb_multi_llrr_split$lri2[duplicated(scdb_multi_llrr_split$lri2) == TRUE] <- ""
scdb_multi_llrr_split$lri3[duplicated(scdb_multi_llrr_split$lri3) == TRUE] <- ""
scdb_multi_llrr_split$lri4[duplicated(scdb_multi_llrr_split$lri4) == TRUE] <- ""
# there are matches between the cols but they are <10

table(is.na(match(scdb_multi_llrr_split$lri1, 
                  lrdb$interactions$interaction_pair)))
table(is.na(match(scdb_multi_llrr_split$lri2, 
                  lrdb$interactions$interaction_pair)))
table(is.na(match(scdb_multi_llrr_split$lri3, 
                  lrdb$interactions$interaction_pair)))
table(is.na(match(scdb_multi_llrr_split$lri4, 
                  lrdb$interactions$interaction_pair)))
# of each type around 10 interactions are not found in lrdb

```

```{r split_lrr, message = FALSE}

# split scdb_multi_lrr
scdb_multi_lrr
scdb_multi_lrr_split <- separate_rows(scdb_multi_lrr, LRI, sep = "_")
scdb_multi_lrr_split$LRI[-grep("&", scdb_multi_lrr_split$LRI)] <- paste(
  scdb_multi_lrr_split$LIGAND_1[-grep("&", scdb_multi_lrr_split$LRI)], 
  scdb_multi_lrr_split$LRI[-grep("&", scdb_multi_lrr_split$LRI)], sep = "&")

table(duplicated(scdb_multi_lrr_split$LRI))
scdb_multi_lrr_split <- scdb_multi_lrr_split[duplicated(
  scdb_multi_lrr_split$LRI) == FALSE,]
table(is.na(match(scdb_multi_lrr_split$LRI, 
                  lrdb$interactions$interaction_pair)))
# actually, the majority of this is already found in lrdb,
# only 140 additional  unique interactions in scdb

scdb_multi_lrr_split[is.na(match(scdb_multi_lrr_split$LRI, 
                                 lrdb$interactions$interaction_pair)),]
# lots of collagens, integrins, and wnt pathway members

# compare to insp_ref
# compare to insp_ref_shared
scdb_multi_lrr_split$ligand_id <- mgi_conv$NCBI.Gene.strand[
  match(scdb_multi_lrr_split$LIGAND_1, mgi_conv$MGI.Accession.ID)]

scdb_multi_lrr_split$receptor1_id <- mgi_conv$NCBI.Gene.strand[
  match(scdb_multi_lrr_split$RECEPTOR_1, mgi_conv$MGI.Accession.ID)]

scdb_multi_lrr_split$receptor2_id <- mgi_conv$NCBI.Gene.strand[
  match(scdb_multi_lrr_split$RECEPTOR_2, mgi_conv$MGI.Accession.ID)]

table(is.na(match(unique(scdb_multi_lrr_split$ligand_id), 
                  insp_ref_shared$ids$ensembl_mbl6_id)))
table(is.na(match(unique(scdb_multi_lrr_split$receptor1_id), 
                  insp_ref_shared$ids$ensembl_mbl6_id)))
table(is.na(match(unique(scdb_multi_lrr_split$receptor2_id), 
                  insp_ref_shared$ids$ensembl_mbl6_id)))
# a majority is in insp_ref_shared

pos_lig_shared <- which(!is.na(match(unique(scdb_multi_lrr_split$ligand_id), 
                  insp_ref_shared$ids$ensembl_mbl6_id)))
pos_rec1_shared <- which(!is.na(match(unique(scdb_multi_lrr_split$receptor1_id), 
                  insp_ref_shared$ids$ensembl_mbl6_id)))
pos_rec2_shared <- which(!is.na(match(unique(scdb_multi_lrr_split$receptor2_id), 
                  insp_ref_shared$ids$ensembl_mbl6_id)))

table(is.na(match(pos_lig_shared, pos_rec1_shared)))
table(is.na(match(pos_lig_shared, pos_rec2_shared)))
# but only around 60 of LR combinations are both found in insp_ref_shared

```

I decided not to add any further interactions, since the benefit is not
worth the additional effort and time.

# Add IDs

For many ligands and receptors ensembl IDs and symbols are missing.
Make use of mgi_conv and hgnc_conv

Technically not needed but more convenient:

```{r load_lrdb2, message = FALSE}

lrdb <- list.load(file = "../../data/databases/lrdb_003.rds")

```

## Ligands

### ENSEMBL IDs

```{r add_ensid_ligands, message = FALSE}

# ensembl IDs stored in: mgi_conv$NCBI.Gene.strand
# symbols stored in: MGI.Accession.ID
# HGNC IDs stored in: CCDS.IDs

head(lrdb$ligands[lrdb$ligands$ligand_ensembl_id == "",])
missing_l <- lrdb$ligands$ligand_symbol[lrdb$ligands$ligand_ensembl_id == ""]
missing_l_uni <- missing_l[duplicated(missing_l) == FALSE]
missing_l_pos <- which(lrdb$ligands$ligand_ensembl_id == "")
length(missing_l_uni)
# around 550 unique ligands do not have ensembl id

table(is.na(match(missing_l_uni, mgi_conv$MGI.Accession.ID)))
table(is.na(match(mgi_conv$MGI.Accession.ID, missing_l_uni)))
# around 500 ensembl IDs can be added

# get the corresponding positions from mgi_conv
mgi_l_pos_uni <- which(!is.na(match(mgi_conv$MGI.Accession.ID, missing_l_uni)))
head(mgi_conv[mgi_l_pos,])
length(mgi_l_pos)
# check if worked
unique(lrdb$ligands$ligand_ensembl_id[missing_l_pos])

# get number match to sort mgi according to lrdb
sort_mgi_l <- match(lrdb$ligands$ligand_symbol[missing_l_pos],
                    mgi_conv$MGI.Accession.ID[mgi_l_pos_uni])
table(is.na(sort_mgi_l))

lrdb$ligands$ligand_ensembl_id[missing_l_pos] <- mgi_conv$NCBI.Gene.strand[
  mgi_l_pos_uni][sort_mgi_l]
head(lrdb$ligands[missing_l_pos,])

```

```{r remove_lig_nas, message = FALSE}

# improve_ only NAs
unique(lrdb$ligands$ligand_ensembl_id[-grep("ENSMUS", 
                                            lrdb$ligands$ligand_ensembl_id)])
lrdb$ligands$ligand_ensembl_id[
  lrdb$ligands$ligand_ensembl_id == "null"] <- NA
table(is.na(lrdb$ligands$ligand_ensembl_id))
# 47 rows of ligands  (23 ligands total) do not have an ensembl ID
na_ligands <- unique(lrdb$ligands$ligand_symbol[
  is.na(lrdb$ligands$ligand_ensembl_id)])
length(na_ligands)

lrdb$interactions <- lrdb$interactions[is.na(match(lrdb$ligands$ligand_symbol,
                                                   na_ligands )),]
lrdb$sources <- lrdb$sources[is.na(match(lrdb$ligands$ligand_symbol,
                                         na_ligands )),]
lrdb$receptors <- lrdb$receptors[is.na(match(lrdb$ligands$ligand_symbol,
                                             na_ligands )),]
lrdb$ligands <- lrdb$ligands[is.na(match(lrdb$ligands$ligand_symbol,
                                         na_ligands )),]

unique(lrdb$ligands$ligand_ensembl_id[-grep("ENSMUS", 
                                            lrdb$ligands$ligand_ensembl_id)])
table(is.na(lrdb$ligands$ligand_ensembl_id))
# no more NAs

```

### HGNC info

```{r add_hgnc_ligands, message = FALSE}

# first hgnc ids
mgi_l_pos <- which(!is.na(match(mgi_conv$NCBI.Gene.strand, 
                                lrdb$ligands$ligand_ensembl_id)))
table(is.na(match(mgi_conv$NCBI.Gene.strand, lrdb$ligands$ligand_ensembl_id)))
length(mgi_l_pos)
length(unique(lrdb$ligands$ligand_ensembl_id))

sort_mgi_l2 <- match(lrdb$ligands$ligand_ensembl_id,
                    mgi_conv$NCBI.Gene.strand[mgi_l_pos])

lrdb$ligands$ligand_hgnc_id <- mgi_conv$CCDS.IDs[mgi_l_pos][sort_mgi_l2]

# then hgnc symbols
hgnc_l_pos <- which(!is.na(match(hgnc_conv$hgnc_id, 
                                 lrdb$ligands$ligand_hgnc_id)))
table(is.na(match(hgnc_conv$hgnc_id, lrdb$ligands$ligand_hgnc_id)))
length(hgnc_l_pos)
length(unique(lrdb$ligands$ligand_hgnc_id))

sort_hgnc_l <- match(lrdb$ligands$ligand_hgnc_id,
                    hgnc_conv$hgnc_id[hgnc_l_pos])

lrdb$ligands$ligand_hgnc_symbol_h <- hgnc_conv$symbol[hgnc_l_pos][sort_hgnc_l]
head(lrdb$ligands)

lrdb$ligands$ligand_hgnc_id[lrdb$ligands$ligand_hgnc_id == "null"] <- NA
table(is.na(lrdb$ligands$ligand_hgnc_id))
table(is.na(lrdb$ligands$ligand_hgnc_symbol_h))

```

## Receptors

### ENSEMBL IDs

```{r add_ensid_ligands, message = FALSE}

head(lrdb$receptors[lrdb$receptors$receptor_ensembl_id == "",])
missing_r <- lrdb$receptors$receptor_symbol[
  lrdb$receptors$receptor_ensembl_id == ""]
missing_r_uni <- missing_r[duplicated(missing_r) == FALSE]
missing_r_pos <- which(lrdb$receptors$receptor_ensembl_id == "")
length(missing_r_uni)
# around 500 unique receptors do not have ensembl id

table(is.na(match(missing_r_uni, mgi_conv$MGI.Accession.ID)))
table(is.na(match(mgi_conv$MGI.Accession.ID, missing_r_uni)))
# around 460 ensembl IDs can be added

# get the corresponding positions from mgi_conv
mgi_r_pos_uni <- which(!is.na(match(mgi_conv$MGI.Accession.ID, missing_r_uni)))
head(mgi_conv[mgi_r_pos,])
length(mgi_r_pos)
# check if worked
unique(lrdb$receptors$receptor_ensembl_id[missing_r_pos])

# get number match to sort mgi according to lrdb
sort_mgi_r <- match(lrdb$receptors$receptor_symbol[missing_r_pos],
                    mgi_conv$MGI.Accession.ID[mgi_r_pos_uni])
table(is.na(sort_mgi_r))

lrdb$receptors$receptor_ensembl_id[missing_r_pos] <- mgi_conv$NCBI.Gene.strand[
  mgi_r_pos_uni][sort_mgi_r]
head(lrdb$receptors[missing_r_pos,])

```

```{r remove_rec_nas, message = FALSE}

# improve_ only NAs
unique(lrdb$receptors$receptor_ensembl_id[-grep(
  "ENSMUS", lrdb$receptors$receptor_ensembl_id)])
table(is.na(lrdb$receptors$receptor_ensembl_id))
# 27 rows of receptorsdo not have an ensembl ID (20 receptors total)
na_receptors <- unique(lrdb$receptors$receptor_symbol[
  is.na(lrdb$receptors$receptor_ensembl_id)])
length(na_receptors)

lrdb$interactions <- lrdb$interactions[is.na(match(
  lrdb$receptors$receptor_symbol, na_receptors )),]
lrdb$sources <- lrdb$sources[is.na(match(lrdb$receptors$receptor_symbol, 
                                         na_receptors )),]
lrdb$ligands <- lrdb$ligands[is.na(match(lrdb$receptors$receptor_symbol,
                                         na_receptors )),]
lrdb$receptors <- lrdb$receptors[is.na(match(lrdb$receptors$receptor_symbol, 
                                             na_receptors )),]

unique(lrdb$receptors$receptor_ensembl_id[-grep("ENSMUS", 
                                            rdb$receptors$receptor_ensembl_id)])
table(is.na(lrdb$receptors$receptor_ensembl_id))
# no more NAs

```

```{r add_hgnc_receptors message = FALSE}

# first hgnc ids
mgi_r_pos <- which(!is.na(match(mgi_conv$NCBI.Gene.strand, 
                                lrdb$receptors$receptor_ensembl_id)))
table(is.na(match(mgi_conv$NCBI.Gene.strand, 
                  lrdb$receptors$receptor_ensembl_id)))
length(mgi_r_pos)
length(unique(lrdb$receptors$receptor_ensembl_id))
# around 800 hgnc ids can be added

sort_mgi_r2 <- match(lrdb$receptors$receptor_ensembl_id,
                    mgi_conv$NCBI.Gene.strand[mgi_r_pos])

lrdb$receptors$receptor_hgnc_id <- mgi_conv$CCDS.IDs[mgi_r_pos][sort_mgi_r2]

# then hgnc symbols
hgnc_r_pos <- which(!is.na(match(hgnc_conv$hgnc_id, 
                                 lrdb$receptors$receptor_hgnc_id)))
table(is.na(match(hgnc_conv$hgnc_id, lrdb$receptors$receptor_hgnc_id)))
length(hgnc_r_pos)
length(unique(lrdb$receptors$receptor_hgnc_id))
#around 700-800 HGNC symbols can be added

sort_hgnc_r <- match(lrdb$receptors$receptor_hgnc_id,
                    hgnc_conv$hgnc_id[hgnc_r_pos])

lrdb$receptors$receptor_hgnc_symbol_h <- hgnc_conv$symbol[
  hgnc_r_pos][sort_hgnc_r]
head(lrdb$receptors)

lrdb$receptors$receptor_hgnc_id[lrdb$receptors$receptor_hgnc_id == "null"] <- NA
table(is.na(lrdb$receptors$receptor_hgnc_id))
table(is.na(lrdb$receptors$receptor_hgnc_id))

```

NAs in HGNC symbols are not so bad.

## Add Info to Rest

```{r add_info, message = FALSE}

lrdb$interactions$ligand_symbol <- lrdb$ligands$ligand_symbol
lrdb$interactions$ligand_ensembl_id <- lrdb$ligands$ligand_ensembl_id

lrdb$interactions$receptor_symbol <- lrdb$receptors$receptor_symbol
lrdb$interactions$receptor_ensembl_id <- lrdb$receptors$receptor_ensembl_id

tail(lrdb$interactions)
tail(lrdb$sources)
tail(lrdb$ligands)
tail(lrdb$receptors)

```

```{r checkpoint3_save, message = FALSE}

list.save(lrdb, file = "../../data/databases/lrdb_004.rds")

```

```{r checkpoint3_load, message = FALSE}

lrdb <- list.load(file = "../../data/databases/lrdb_004.rds")

```

# Compare to insp_ref_shared

```{r comp_insp_ref_shared, message = FALSE}

# number of matching ligands/receptors
table(is.na(match(unique(lrdb$ligands$ligand_ensembl_id),
                  insp_ref_shared$ids$ensembl_mbl6_id)))
table(is.na(lrdb$ligands$ligand_ensembl_id))
table(duplicated(lrdb$ligands$ligand_ensembl_id))
# most ligands are duplicated

table(is.na(match(unique(lrdb$receptors$receptor_ensembl_id), 
                  insp_ref_shared$ids$ensembl_mbl6_id)))

# number of matching rows
table(is.na(match(lrdb$ligands$ligand_ensembl_id,
                  insp_ref_shared$ids$ensembl_mbl6_id)))
table(is.na(match(lrdb$receptors$receptor_ensembl_id,
                  insp_ref_shared$ids$ensembl_mbl6_id)))

```

## Ligands

```{r ligands_tb_discarded, message = FALSE}

lrdb$ligands[(is.na(match(lrdb$ligands$ligand_ensembl_id, 
                          insp_ref_shared$ids$ensembl_mbl6_id))),]
l_tbd <- unique(lrdb$ligands$ligand_symbol[
  (is.na(match(lrdb$ligands$ligand_ensembl_id, 
               insp_ref_shared$ids$ensembl_mbl6_id)))])
length(l_tbd)
l_tbd
# 122 receptors to be removed

```

```{r discard_ligands_rows, message = FALSE}

# get the rows of ligands to discard

l_tbd_pos <- which(!is.na(match(lrdb$ligands$ligand_symbol, l_tbd)))
length(l_tbd_pos)
lrdb$interactions[l_tbd_pos,]

# 300 rows to be removed
lrdb$interactions <- lrdb$interactions[-l_tbd_pos,]
lrdb$sources <- lrdb$sources[-l_tbd_pos,]
lrdb$ligands <- lrdb$ligands[-l_tbd_pos,]
lrdb$receptors <- lrdb$receptors[-l_tbd_pos,]

table(is.na(match(lrdb$ligands$ligand_ensembl_id,
                  insp_ref_shared$ids$ensembl_mbl6_id)))

```

## Receptors

```{r receptors_tb_discarded, message = FALSE}

lrdb$receptors[(is.na(match(lrdb$receptors$receptor_ensembl_id,
                            insp_ref_shared$ids$ensembl_mbl6_id))),]
r_tbd <- unique(lrdb$receptors$receptor_symbol[
  (is.na(match(lrdb$receptors$receptor_ensembl_id, 
               insp_ref_shared$ids$ensembl_mbl6_id)))])
length(r_tbd)
r_tbd
# 71 receptors to be discarded
# looks like there is some overlap with l_tbd

table(is.na(match(l_tbd, r_tbd)))
# just a little

```

```{r discard_receptors_rows, message = FALSE}

# get the rows of ligands to discard

r_tbd_pos <- which(!is.na(match(lrdb$receptors$receptor_symbol, r_tbd)))
length(r_tbd_pos)
lrdb$interactions[r_tbd_pos,]

# 300 rows to be removed
lrdb$interactions <- lrdb$interactions[-r_tbd_pos,]
lrdb$sources <- lrdb$sources[-r_tbd_pos,]
lrdb$ligands <- lrdb$ligands[-r_tbd_pos,]
lrdb$receptors <- lrdb$receptors[-r_tbd_pos,]

table(is.na(match(lrdb$ligands$ligand_ensembl_id,
                  insp_ref_shared$ids$ensembl_mbl6_id)))

#  around 2400 interactions that are also found in insp_ref are left
tail(lrdb$interactions)
tail(lrdb$sources)
tail(lrdb$ligands)
tail(lrdb$receptors)

```


```{r checkpoint3_save, message = FALSE}

list.save(lrdb, file = "../../data/databases/lrdb_005.rds")

```

```{r checkpoint3_load, message = FALSE}

lrdb <- list.load(file = "../../data/databases/lrdb_005.rds")

```

# Add Annotations

Around 380 interactions do not have functional annotations.

```{r check_anno, message = FALSE}

# number of missing annotations
nrow(lrdb$interactions[lrdb$interactions$interaction_type == "",])
lrdb$interactions[lrdb$interactions$interaction_type == "",]
no_int_type <- lrdb$interactions$interaction_pair[
  lrdb$interactions$interaction_type == ""]

# number of conflicting annotations
table(lrdb$interaction$interaction_type[
  lrdb$interaction$interaction_type != ""])

conflict_pos <- c(
  grep("Cell-Cell Contact,Secreted", lrdb$interaction$interaction_type),
  grep("ECM-Receptor,Secreted", lrdb$interaction$interaction_type),
  grep("Secreted Signaling,Membrane", lrdb$interaction$interaction_type),
  grep("Secreted Signaling,ECM", lrdb$interaction$interaction_type)
 )
length(conflict_pos)
lrdb$interaction$interaction_pair[conflict_pos]
# 36 conflicting annotations left

```


