---
author: "lea wölbert"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preparation

## Libraries

```{r libraries, message = FALSE}

library(SingleCellExperiment)
library(tidyverse)
library(scuttle)
library(scran)
library(rlist)

```

## Load

### Code

```{r load_code, message = FALSE}

source(file = "../code/ext_code/InteractionRanking.r")
source(file = "../code/source_code/cci_functions.r")
source(file = "../code/source_code/colourscheme.R")
source(file = "../code/source_code/filepath.R")

```

### Objects

```{r load_lrdb, message = FALSE}

lrdb <- list.load(file = paste(
  filepath_lv2,
  "data/databases/lrdb_anno.rds",
  sep = "")
)

lrdb_part <- list.load(file = paste(
  filepath_lv2,
  "data/databases/lrdb_part.rds",
  sep = "")
)

```

```{r load_mgi, message = FALSE}

mgi_conv <- read.delim(file = paste(
  filepath_lv2,
  "data/databases/HGNC_AllianceHomology.rpt.txt",
  sep = ""),
  header = TRUE)

```

# Check

Use the mgi conversion table to check if the names match between mgi and lrdb.

```{r}

mgi_conv

test_ligandlist <- mgi_conv$MGI.Accession.ID[
  !is.na(match(mgi_conv$NCBI.Gene.strand, 
               lrdb$interactions$ligand_ensembl_id))]

setequal(test_ligandlist, lrdb$interactions$ligand_symbol)

```

They don't. Why?

```{r}

length(unique(lrdb$interactions$ligand_symbol))
length(unique(lrdb$interactions$ligand_ensembl_id))
length(unique(lrdb$interactions$receptor_symbol))
length(unique(lrdb$interactions$receptor_ensembl_id))
# one ligand ID must have two ligand symbols

ligand_id <- lrdb$interactions$ligand_ensembl_id

for(i in unique(ligand_id)){
  print(i)
  print(lrdb$interactions$ligand_symbol[
    lrdb$interactions$ligand_ensembl_id == i][
      duplicated(lrdb$interactions$ligand_symbol[
    lrdb$interactions$ligand_ensembl_id == i]) == FALSE
    ])
}

# the culprit is ENSMUSG00000048534
lrdb$interactions[lrdb$interactions$ligand_ensembl_id == "ENSMUSG00000048534",]
lrdb$interactions[lrdb$interactions$ligand_symbol == "Amica1",]
lrdb$sources[lrdb$interactions$ligand_symbol == "Amica1",]

# JAML1 is the correct gene name

# remove Amica1
lrdb$sources <- lrdb$sources[lrdb$interactions$ligand_symbol != "Amica1",]
lrdb$ligands <- lrdb$ligands[lrdb$interactions$ligand_symbol != "Amica1",]
lrdb$receptors <- lrdb$receptors[lrdb$interactions$ligand_symbol != "Amica1",]
lrdb$interactions <- lrdb$interactions[
  lrdb$interactions$ligand_symbol != "Amica1",]

nrow(lrdb$sources)

length(unique(lrdb$interactions$ligand_symbol))
length(unique(lrdb$interactions$ligand_ensembl_id))
length(unique(lrdb$interactions$receptor_symbol))
length(unique(lrdb$interactions$receptor_ensembl_id))

```

## Test ligands

Check if there ligands all match mgi.
The remaining non-matching gene names are most likely "old" gene names
just taken from mjdb that were updated in mgi in the meantime.

```{r}

test_ligandlist <- mgi_conv$MGI.Accession.ID[
  !is.na(match(mgi_conv$NCBI.Gene.strand, 
               lrdb$interactions$ligand_ensembl_id))]

setequal(test_ligandlist, unique(lrdb$interactions$ligand_symbol))
# still not identical

length(test_ligandlist)
length(unique(lrdb$interactions$ligand_symbol))
# two IDs too many

lrdb$interactions[is.na(match(
  lrdb$interactions$ligand_symbol, test_ligandlist)),]
lrdb$sources[is.na(match(
  lrdb$interactions$ligand_symbol, test_ligandlist)),]

# get all ligands that don't match
test_ligandlist[is.na(match(
  test_ligandlist, unique(lrdb$interactions$ligand_symbol)
))]

```

```{r}

# find outlier ligands
mgi_conv[grep("ENSMUSG00000031380", mgi_conv$NCBI.Gene.strand),]
#Figf
#Vegfd

mgi_conv[grep("ENSMUSG00000037362", mgi_conv$NCBI.Gene.strand),]
#Nov
#Ccn3

mgi_conv[grep("ENSMUSG00000062991", mgi_conv$NCBI.Gene.strand),]
mgi_conv[grep("ENSMUSG00000118541", mgi_conv$NCBI.Gene.strand),]
# these IDs are not found in mgi_conv
# but the one I found on http://www.ensembl.org is

mgi_conv[grep("ENSMUSG00000100916", mgi_conv$NCBI.Gene.strand),]
mgi_conv[grep("ENSMUSG00000118462", mgi_conv$NCBI.Gene.strand),]
# these IDs are not found in mgi_conv
# but the one I found on http://www.ensembl.org is

# these gene names are indeed not found in lrdb
grep("Vegfd", lrdb$interactions$ligand_symbol)
grep("Vegfd", lrdb$interactions$receptor_symbol)
grep("Ccn3", lrdb$interactions$ligand_symbol)
grep("Ccn3", lrdb$interactions$receptor_symbol)

```

Rename step by step. 
Figf was incorrect, vegf correct.

```{r}

vegf_position <- which(lrdb$interactions$ligand_symbol == "Figf")

lrdb$ligands$ligand_symbol[vegf_position] <- gsub(
  "Figf", "Vegfd", lrdb$ligands$ligand_symbol[vegf_position])

lrdb$interactions$interaction_pair[vegf_position] <- gsub(
  "Figf", "Vegfd", lrdb$interactions$interaction_pair[vegf_position])

lrdb$interactions$ligand_symbol[vegf_position] <- gsub(
  "Figf", "Vegfd", lrdb$interactions$ligand_symbol[vegf_position])

lrdb$sources$interaction_pair[vegf_position] <- gsub(
  "Figf", "Vegfd", lrdb$sources$interaction_pair[vegf_position])

lrdb$interactions[vegf_position,]

```

Nov was incorrect, Ccn3 correct.

```{r}

ccn3_position <- which(lrdb$interactions$ligand_symbol == "Nov")

lrdb$ligands$ligand_symbol[ccn3_position] <- gsub(
  "Nov", "Ccn3", lrdb$ligands$ligand_symbol[ccn3_position])

lrdb$interactions$interaction_pair[ccn3_position] <- gsub(
  "Nov", "Ccn3", lrdb$interactions$interaction_pair[ccn3_position])

lrdb$interactions$ligand_symbol[ccn3_position] <- gsub(
  "Nov", "Ccn3", lrdb$interactions$ligand_symbol[ccn3_position])

lrdb$sources$interaction_pair[ccn3_position] <- gsub(
  "Nov", "Ccn3", lrdb$sources$interaction_pair[ccn3_position])

lrdb$interactions[ccn3_position,]

```

Also, in two cases the IDs were deprecated.
Since the "old" IDs from mjdb are also the ones found in the SCE 
objects/insp_ref, I decided to keep those two IDs or it will lead to problems
later.

```{r}

# will not change the IDs because the "old" Ids in LRDB are also in insp_ref,
# so if I change, these genes will be lost from the SCE objects

#lhb_position <- which(
#lrdb$interactions$ligand_ensembl_id == "ENSMUSG00000100916")

#lrdb$ligands$ligand_ensembl_id[lhb_position] <- gsub("ENSMUSG00000100916", 
#"ENSMUSG00000118462", lrdb$ligands$ligand_ensembl_id[lhb_position])


#lrdb$interactions$ligand_ensembl_id[lhb_position] <- gsub(
#"ENSMUSG00000100916", "ENSMUSG00000118462", 
#lrdb$interactions$ligand_ensembl_id[lhb_position])

#lrdb$interactions[lhb_position,]

```

```{r}

# will not change the IDs because the "old" Ids in LRDB are also in insp_ref,
# so if I change, these genes will be lost from the SCE objects

#ngr1_position <- which(l
#rdb$interactions$ligand_ensembl_id == "ENSMUSG00000062991")

#lrdb$ligands$ligand_ensembl_id[ngr1_position] <- gsub(
#"ENSMUSG00000062991", "ENSMUSG00000118541", 
#lrdb$ligands$ligand_ensembl_id[ngr1_position])


#lrdb$interactions$ligand_ensembl_id[ngr1_position] <- gsub(
#"ENSMUSG00000062991", "ENSMUSG00000118541", 
#lrdb$interactions$ligand_ensembl_id[ngr1_position])

#lrdb$interactions[ngr1_position,]

```

Test again if they match.

```{r}

test_ligandlist <- mgi_conv$MGI.Accession.ID[
  !is.na(match(mgi_conv$NCBI.Gene.strand, 
               lrdb$interactions$ligand_ensembl_id))]


setequal(test_ligandlist, unique(lrdb$interactions$ligand_symbol))

lrdb$interactions[is.na(match(
  lrdb$interactions$ligand_symbol, test_ligandlist)),]
lrdb$sources[is.na(match(
  lrdb$interactions$ligand_symbol, test_ligandlist)),]

test_ligandlist[is.na(match(
  test_ligandlist, unique(lrdb$interactions$ligand_symbol)
))]

```

Still no perfect match but the only remaining ligand IDs are the ones that I 
decided to keep.

# Receptor list

Repeat for receptors.

```{r}

test_receptorlist <- mgi_conv$MGI.Accession.ID[
  !is.na(match(mgi_conv$NCBI.Gene.strand, 
               lrdb$interactions$receptor_ensembl_id))]


setequal(test_receptorlist, unique(lrdb$interactions$receptor_symbol))

# still not identical
length(test_receptorlist)
length(unique(lrdb$interactions$receptor_symbol))

lrdb$interactions[is.na(match(
  lrdb$interactions$receptor_symbol, test_receptorlist)),]
lrdb$sources[is.na(match(
  lrdb$interactions$receptor_symbol, test_receptorlist)),]

test_receptorlist[is.na(match(
  test_receptorlist, lrdb$interactions$receptor_symbol
))]

```

A few more receptors must be updated step by step.

```{r}

mgi_conv[grep("ENSMUSG00000037362", mgi_conv$NCBI.Gene.strand),]
# nov
# ccn3

mgi_conv[grep("ENSMUSG00000006378", mgi_conv$NCBI.Gene.strand),]
# Galr3
# Gcat

mgi_conv[grep("ENSMUSG00000002885", mgi_conv$NCBI.Gene.strand),]
# Cd97
# Adgre5

mgi_conv[grep("ENSMUSG00000038403", mgi_conv$NCBI.Gene.strand),]
# Hfe2
# Hjv

mgi_conv[grep("ENSMUSG00000031785", mgi_conv$NCBI.Gene.strand),]
# Gpr56
# Adgrg1

mgi_conv[grep("ENSMUSG00000046432", mgi_conv$NCBI.Gene.strand),]
# Ngfrap1
# Bex3

```

```{r}

galr_position <- which(lrdb$interactions$receptor_symbol == "Galr3")

lrdb$receptors$receptor_symbol[galr_position] <- gsub(
  "Galr3", "Gcat", lrdb$receptors$receptor_symbol[galr_position])

lrdb$interactions$interaction_pair[galr_position] <- gsub(
  "Galr3", "Gcat", lrdb$interactions$interaction_pair[galr_position])

lrdb$interactions$receptor_symbol[galr_position] <- gsub(
  "Galr3", "Gcat", lrdb$interactions$receptor_symbol[galr_position])

lrdb$sources$interaction_pair[galr_position] <- gsub(
  "Galr3", "Gcat", lrdb$sources$interaction_pair[galr_position])

lrdb$interactions[galr_position,]
lrdb$receptors[galr_position,]

```

```{r}

ccn3_position <- which(lrdb$interactions$receptor_symbol == "Nov")

lrdb$receptors$receptor_symbol[ccn3_position] <- gsub(
  "Nov", "Ccn3", lrdb$receptors$receptor_symbol[ccn3_position])

lrdb$interactions$interaction_pair[ccn3_position] <- gsub(
  "Nov", "Ccn3", lrdb$interactions$interaction_pair[ccn3_position])

lrdb$interactions$receptor_symbol[ccn3_position] <- gsub(
  "Nov", "Ccn3", lrdb$interactions$receptor_symbol[ccn3_position])

lrdb$sources$interaction_pair[ccn3_position] <- gsub(
  "Nov", "Ccn3", lrdb$sources$interaction_pair[ccn3_position])

lrdb$interactions[ccn3_position,]

```

```{r}

cd97_position <- which(lrdb$interactions$receptor_symbol == "Cd97")

lrdb$receptors$receptor_symbol[cd97_position] <- gsub(
  "Cd97", "Adgre5", lrdb$receptors$receptor_symbol[cd97_position])

lrdb$interactions$interaction_pair[cd97_position] <- gsub(
  "Cd97", "Adgre5", lrdb$interactions$interaction_pair[cd97_position])

lrdb$interactions$receptor_symbol[cd97_position] <- gsub(
  "Cd97", "Adgre5", lrdb$interactions$receptor_symbol[cd97_position])

lrdb$sources$interaction_pair[cd97_position] <- gsub(
  "Cd97", "Adgre5", lrdb$sources$interaction_pair[cd97_position])

lrdb$interactions[cd97_position,]
lrdb$receptors[cd97_position,]

```

```{r}

hfe2_position <- which(lrdb$interactions$receptor_symbol == "Hfe2")

lrdb$receptors$receptor_symbol[hfe2_position] <- gsub(
  "Hfe2", "Hjv", lrdb$receptors$receptor_symbol[hfe2_position])

lrdb$interactions$interaction_pair[hfe2_position] <- gsub(
  "Hfe2", "Hjv", lrdb$interactions$interaction_pair[hfe2_position])

lrdb$interactions$receptor_symbol[hfe2_position] <- gsub(
  "Hfe2", "Hjv", lrdb$interactions$receptor_symbol[hfe2_position])

lrdb$sources$interaction_pair[hfe2_position] <- gsub(
  "Hfe2", "Hjv", lrdb$sources$interaction_pair[hfe2_position])

lrdb$interactions[hfe2_position,]
lrdb$receptors[hfe2_position,]

```

```{r}

gpr56_position <- which(lrdb$interactions$receptor_symbol == "Gpr56")

lrdb$receptors$receptor_symbol[gpr56_position] <- gsub(
  "Gpr56", "Adgrg1", lrdb$receptors$receptor_symbol[gpr56_position])

lrdb$interactions$interaction_pair[gpr56_position] <- gsub(
  "Gpr56", "Adgrg1", lrdb$interactions$interaction_pair[gpr56_position])

lrdb$interactions$receptor_symbol[gpr56_position] <- gsub(
  "Gpr56", "Adgrg1", lrdb$interactions$receptor_symbol[gpr56_position])

lrdb$sources$interaction_pair[gpr56_position] <- gsub(
  "Gpr56", "Adgrg1", lrdb$sources$interaction_pair[gpr56_position])

lrdb$interactions[gpr56_position,]
lrdb$receptors[gpr56_position,]

```

```{r}

ngfrap1_position <- which(lrdb$interactions$receptor_symbol == "Ngfrap1")

lrdb$receptors$receptor_symbol[ngfrap1_position] <- gsub(
  "Ngfrap1", "Bex3", lrdb$receptors$receptor_symbol[ngfrap1_position])

lrdb$interactions$interaction_pair[ngfrap1_position] <- gsub(
  "Ngfrap1", "Bex3", lrdb$interactions$interaction_pair[ngfrap1_position])

lrdb$interactions$receptor_symbol[ngfrap1_position] <- gsub(
  "Ngfrap1", "Bex3", lrdb$interactions$receptor_symbol[ngfrap1_position])

lrdb$sources$interaction_pair[ngfrap1_position] <- gsub(
  "Ngfrap1", "Bex3", lrdb$sources$interaction_pair[ngfrap1_position])

lrdb$interactions[ngfrap1_position,]
lrdb$receptors[ngfrap1_position,]

```

Check if now everything matches to mgi.

```{r}

test_receptorlist <- mgi_conv$MGI.Accession.ID[
  !is.na(match(mgi_conv$NCBI.Gene.strand, 
               lrdb$interactions$receptor_ensembl_id))]

setequal(test_receptorlist, unique(lrdb$interactions$receptor_symbol))

```

It does.

```{r save_lrdb_comp, message = FALSE}

list.save(lrdb, file = paste(
  filepath_lv2,
  "data/databases/lrdb_comp.rds",
  sep = "")
)

```

Save under different name to account for all manual changes.
Now LRDB_COMP is the lrdb to be used for CCI analysis.

# lrdb_part

Make lrdb_part to use in same format as mjdb.

```{r save_lrdb_part, message = FALSE}

lrdb_part <- data.frame(
  Receptor_ENSEMBL = lrdb$receptors$receptor_ensembl_id,
  Ligand_ENSEMBL = lrdb$ligands$ligand_ensembl_id,
  Receptor_GeneSymbol = lrdb$receptors$receptor_symbol,
  Ligand_GeneSymbol = lrdb$ligands$ligand_symbol
)

list.save(lrdb_part, file = paste(
  filepath_lv2,
  "data/databases/lrdb_part.rds",
  sep = "")
)

```
