---
author: "perrine lacour lea woelbert"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(biomaRt)
```

This part was taken from Perrine's script on how to download the relevant data 
from BioMart.

```{r load, message = FALSE}
ensembl_mouse <- useMart(biomart="ENSEMBL_MART_ENSEMBL",
                         host="www.ensembl.org",
                         dataset="mmusculus_gene_ensembl")
ensembl_list_mbl6 <- getBM(attributes=c("ensembl_gene_id",
                                        "external_gene_name"), 
                           mart=ensembl_mouse)

ensembl_mcar <- useMart(biomart="ENSEMBL_MART_ENSEMBL",
                        host="www.ensembl.org",
                        dataset="mcaroli_gene_ensembl")
ensembl_list_mcar <- getBM(attributes=c("ensembl_gene_id",
                                        "mmusculus_homolog_ensembl_gene",
                                      "mmusculus_homolog_associated_gene_name"), 
                           mart=ensembl_mcar)

ensembl_mcas <- useMart(biomart="ENSEMBL_MART_MOUSE",
                        host="www.ensembl.org",
                        dataset="mmcasteij_gene_ensembl")
ensembl_list_mcas <- getBM(attributes=c("ensembl_gene_id",
                                        "mmusculus_homolog_ensembl_gene",
                                      "mmusculus_homolog_associated_gene_name"), 
                           mart=ensembl_mcas)

ensembl_mspr <- useMart(biomart="ENSEMBL_MART_ENSEMBL",
                        host="www.ensembl.org",
                        dataset="mspretus_gene_ensembl")
ensembl_list_mspr=getBM(attributes=c("ensembl_gene_id",
                                     "mmusculus_homolog_ensembl_gene",
                                     "mmusculus_homolog_associated_gene_name"), 
                        mart=ensembl_mspr)

```

```{r look_at_mouse, message = FALSE}

dim(ensembl_list_mbl6)
head(ensembl_list_mbl6)

# how many ensembl_gene_id are duplicated empty?
table(duplicated(ensembl_list_mbl6$ensembl_gene_id)) 
table(ensembl_list_mbl6$ensembl_gene_id == "")

# how many external.gene.names are duplicated empty?
table(duplicated(ensembl_list_mbl6$external_gene_name)) 
table(ensembl_list_mbl6$external_gene_name == "")

# look at duplicated external gene names
# which of the duplicated external gene names ones are empty?
dup_ensembl_list_mbl6 <- ensembl_list_mbl6[duplicated(
  ensembl_list_mbl6$external_gene_name),]
table(dup_ensembl_list_mbl6$external_gene_name == "")

empty_ensembl_list_mbl6<- dup_ensembl_list_mbl6[
  dup_ensembl_list_mbl6$external_gene_name == "",]
nempty_ensembl_list_mbl6<- dup_ensembl_list_mbl6[
  dup_ensembl_list_mbl6$external_gene_name != "",]


# How many times were IDs duplicated (= distribution of mapping number)
# Taken from Perrine
table(table(ensembl_list_mbl6$ensembl_gene_id))
# in this case all were mapped exactly once

# How many times were external_gene_names duplicated 
table(table(ensembl_list_mbl6$external_gene_name))
# in this case all were mapped exactly once

```

There are no duplicated or empty ensembl IDs for mouse.
There are almost 1500 duplicated external_gene_names.
This means that there are multiple IDs for various genes.
ISome external_gene_names were even duplicated over 100 times,
"" was duplicated 1492 times.

```{r look_at_mcar, message = FALSE}

dim(ensembl_list_mcar)
head(ensembl_list_mcar)

# how many ensembl_gene_id are duplicated empty?
table(duplicated(ensembl_list_mcar$ensembl_gene_id)) 
table(ensembl_list_mcar$ensembl_gene_id == "") 

# how many mmusculus_homolog_ensembl_gene are duplicated empty?
table(duplicated(ensembl_list_mcar$mmusculus_homolog_ensembl_gene)) 
table(ensembl_list_mcar$mmusculus_homolog_ensembl_gene == "")

# how many mmusculus_homolog_associated_gene_name are duplicated or empty?
table(duplicated(ensembl_list_mcar$mmusculus_homolog_associated_gene_name)) 
table(ensembl_list_mcar$mmusculus_homolog_associated_gene_name == "")

# take a look at duplicated
# how many not empty mmusculus_homolog_ensembl_genes are duplicated?
table(duplicated(ensembl_list_mcar$mmusculus_homolog_ensembl_gene[
  ensembl_list_mcar$mmusculus_homolog_ensembl_gene != ""]))
# how many not empty mmusculus_homolog_associated_gene_name are duplicated?
table(duplicated(ensembl_list_mcar$mmusculus_homolog_associated_gene_name[
  ensembl_list_mcar$mmusculus_homolog_associated_gene_name != ""]))


# How many times were IDs duplicated (= distribution of mapping number)
# Taken from Perrine
table(table(ensembl_list_mcar$ensembl_gene_id))

# How many times were mmusculus_homolog_ensembl_gene duplicated 
table(table(ensembl_list_mcar$mmusculus_homolog_ensembl_gene))
# How many times were mmusculus_homolog_associated_gene_name duplicated 
table(table(ensembl_list_mcar$mmusculus_homolog_associated_gene_name))

```

The ensembl IDs may be duplicated, but they are not empty.
Some ensembl IDs were duplicated many times.
The number of duplicated gene names is quite high (10.000) but a lot of 
those appear to be empty, only around 500 duplicated genes are left.

```{r look_at_mcas, message = FALSE}

dim(ensembl_list_mcas)
head(ensembl_list_mcas)

# how many ensembl_gene_id are duplicated or empty?
table(duplicated(ensembl_list_mcas$ensembl_gene_id)) 
table(ensembl_list_mcas$ensembl_gene_id == "") 

# how many mmusculus_homolog_ensembl_gene are duplicated or empty?
table(duplicated(ensembl_list_mcas$mmusculus_homolog_ensembl_gene)) 
table(ensembl_list_mcas$mmusculus_homolog_ensembl_gene == "")

# how many mmusculus_homolog_associated_gene_name are duplicated or empty?
table(duplicated(ensembl_list_mcas$mmusculus_homolog_associated_gene_name)) 
table(ensembl_list_mcas$mmusculus_homolog_associated_gene_name == "")

# take a look at duplicated
# how many not empty mmusculus_homolog_associated_gene_name are duplicated?
table(duplicated(ensembl_list_mcas$mmusculus_homolog_ensembl_gene[
  ensembl_list_mcas$mmusculus_homolog_ensembl_gene != ""]))
# how many not empty mmusculus_homolog_associated_gene_name are duplicated?
table(duplicated(ensembl_list_mcas$mmusculus_homolog_associated_gene_name[
  ensembl_list_mcas$mmusculus_homolog_associated_gene_name != ""]))


# How many times were IDs duplicated 
table(table(ensembl_list_mcas$ensembl_gene_id))

# How many times were mmusculus_homolog_ensembl_gene duplicated 
table(table(ensembl_list_mcas$mmusculus_homolog_ensembl_gene))
# How many times were mmusculus_homolog_associated_gene_name duplicated 
table(table(ensembl_list_mcas$mmusculus_homolog_associated_gene_name))

```

Again, few duplicated ensembl_gene_ids but no empty ones.
Interestingly, mcas ensembl_ids were often not mapped multiple times.
The cas mmusculus_homolog_associated_gene_names are overwhelmingly empty, but
there are still around 27725 mmusculus_homolog_ensembl_gene names that are
neither empty nor duplicated.
I guess this means that mcas is the least annotated dataset and that 
mmusculus_homolog_associated_gene_name for mcas doesn't contain that much info.

```{r look_at_mspr, message = FALSE}

dim(ensembl_list_mspr)
head(ensembl_list_mspr)

# how many ensembl_gene_id are duplicated or empty?
table(duplicated(ensembl_list_mspr$ensembl_gene_id)) 
table(ensembl_list_mspr$ensembl_gene_id == "") 

# how many mmusculus_homolog_ensembl_gene are duplicated or empty?
table(duplicated(ensembl_list_mspr$mmusculus_homolog_ensembl_gene)) 
table(ensembl_list_mspr$mmusculus_homolog_ensembl_gene == "")

# how many mmusculus_homolog_associated_gene_name are duplicated or empty?
table(duplicated(ensembl_list_mspr$mmusculus_homolog_associated_gene_name)) 
table(ensembl_list_mspr$mmusculus_homolog_associated_gene_name == "")

# take a look at duplicated
# how many not empty mmusculus_homolog_associated_gene_name are duplicated?
table(duplicated(ensembl_list_mspr$mmusculus_homolog_ensembl_gene[
  ensembl_list_mspr$mmusculus_homolog_ensembl_gene != ""]))
# how many not empty mmusculus_homolog_associated_gene_name are duplicated?
table(duplicated(ensembl_list_mspr$mmusculus_homolog_associated_gene_name[
  ensembl_list_mspr$mmusculus_homolog_associated_gene_name != ""]))


# How many times were IDs duplicated (= distribution of mapping number)
# Taken from Perrine
table(table(ensembl_list_mspr$ensembl_gene_id))

# How many times were mmusculus_homolog_ensembl_gene duplicated 
table(table(ensembl_list_mspr$mmusculus_homolog_ensembl_gene))
# How many times were mmusculus_homolog_associated_gene_name duplicated 
table(table(ensembl_list_mspr$mmusculus_homolog_associated_gene_name))

```

Again, few duplicated ensembl_gene_ids but no empty ones.
For mspr, the number of empty or duplicates gene names is generally not
quite that high similarly to mouse and mcar.
