---
author: "lea wölbert"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, message = FALSE}
library(biomaRt)
library(rlist)
```

This part was taken from Perrine's script on how to download the relevant data 
from BioMart.

```{r load_biomart, message = FALSE}

ensembl_mouse <- useMart(biomart="ENSEMBL_MART_ENSEMBL",
                         host="www.ensembl.org",
                         dataset="mmusculus_gene_ensembl")
ensembl_list_mbl6 <- getBM(attributes=c("ensembl_gene_id",
                                        "external_gene_name"), 
                           mart=ensembl_mouse)

ensembl_mcar <- useMart(biomart="ENSEMBL_MART_ENSEMBL",
                        host="www.ensembl.org",
                        dataset="mcaroli_gene_ensembl")
ensembl_list_mcar <- getBM(attributes=c("ensembl_gene_id",
                                        "mmusculus_homolog_ensembl_gene",
                                      "mmusculus_homolog_associated_gene_name"), 
                           mart=ensembl_mcar)

ensembl_mcas <- useMart(biomart="ENSEMBL_MART_MOUSE",
                        host="www.ensembl.org",
                        dataset="mmcasteij_gene_ensembl")
ensembl_list_mcas <- getBM(attributes=c("ensembl_gene_id",
                                        "mmusculus_homolog_ensembl_gene",
                                      "mmusculus_homolog_associated_gene_name"), 
                           mart=ensembl_mcas)

ensembl_mspr <- useMart(biomart="ENSEMBL_MART_ENSEMBL",
                        host="www.ensembl.org",
                        dataset="mspretus_gene_ensembl")
ensembl_list_mspr <- getBM(attributes=c("ensembl_gene_id",
                                        "mmusculus_homolog_ensembl_gene",
                                      "mmusculus_homolog_associated_gene_name"), 
                        mart=ensembl_mspr)

```

I have an approximate idea of what the data generally looks like from 
01_load_check_data.
Here I will attempt to merge the information by matching the ensembl_gene_ids.
The ensembl_of mbl6 must match the mmusculus_homolog_ensembl_gene of
other species.

This script is used to generate a generate reference dataset containing
all IDs, including one-to-many mappings and duplicated IDs or gene names.
The aim is to add all information and make this dataset as complete as possible.

Later on, this dataset will be subset into more useful datasets, for 
instance containing less duplicates/one-to-one mappings or only genes/IDs 
that are detected in the SCE objects that I will be dealing with.

# Generate Scaffold

First, make an empty scaffold based on the mbl6 ensembl ids.
Every mbl6 ensembl ID appears exactly once, this vector is therefore the
perfect backbone to start a huge DF list which can then be reduced 
according to data requirements and gaps in other data.

```{r make_insp_ref_scaffold, message = FALSE}

# insp_ref = interspecies reference

insp_ref <- list(
  ids = data.frame(
    ensembl_mbl6_id = ensembl_list_mbl6$ensembl_gene_id,
    ensembl_mcar_id = vector("character", length = nrow(ensembl_list_mbl6)),
    ensembl_mcas_id = vector("character", length = nrow(ensembl_list_mbl6)),
    ensembl_mspr_id = vector("character", length = nrow(ensembl_list_mbl6))),
  
  gene_names = data.frame(
    ensembl_mbl6_id = ensembl_list_mbl6$ensembl_gene_id,
    mbl6_external_gene_name = vector(
      "character", length = nrow(ensembl_list_mbl6)),
    mcar_mmusculus_homolog_associated_gene_name = vector(
      "character", length = nrow(ensembl_list_mbl6)),
    mcas_mmusculus_homolog_associated_gene_name = vector(
      "character", length = nrow(ensembl_list_mbl6)),
    mspr_mmusculus_homolog_associated_gene_name = vector(
      "character", length = nrow(ensembl_list_mbl6))),
  
  synonyms = data.frame(
    ensembl_mbl6_id = ensembl_list_mbl6$ensembl_gene_id,
    mbl6_external_gene_name = vector("character", 
                                     length = nrow(ensembl_list_mbl6)),
    synonym.1 = vector("character", length = nrow(ensembl_list_mbl6)),
    synonym.2 = vector("character", length = nrow(ensembl_list_mbl6)),
    synonym.3 = vector("character", length = nrow(ensembl_list_mbl6)),
    synonym.4 = vector("character", length = nrow(ensembl_list_mbl6)),
    synonym.5 = vector("character", length = nrow(ensembl_list_mbl6)),
    synonym.6 = vector("character", length = nrow(ensembl_list_mbl6))),
  
  metadata = data.frame(
    ensembl_mbl6_id = ensembl_list_mbl6$ensembl_gene_id,
    duplicated = vector("character", length = nrow(ensembl_list_mbl6)),
    source = rep("ensembl_mouse", length = nrow(ensembl_list_mbl6)),
    notes = vector("character", length = nrow(ensembl_list_mbl6))
    )
)
                                  
```

# Add Data 

## Easy Matches

### BL6/Mouse

```{r add_bl6, message = FALSE}

# For mbl6, I don't have to match because the order is already identical.
insp_ref$gene_names$mbl6_external_gene_name <- ensembl_list_mbl6$external_gene_name

```

### Caroli

```{r add_mcar_check, message = FALSE}

head(ensembl_list_mcar)
nrow(ensembl_list_mcar)

# insp_ref positions that will be filled with corresponding data from list_mcar
mcar_insp_fill_pos <- match(insp_ref$ids$ensembl_mbl6_id, 
                            ensembl_list_mcar$mmusculus_homolog_ensembl_gene)
# should be same length is insp rows
nrow(insp_ref$ids)
length(mcar_insp_fill_pos)

# positions for ensembl_list_mcar that will replace empty slots in insp_ref
mcar_replace_pos <- match(ensembl_list_mcar$mmusculus_homolog_ensembl_gene,
                          insp_ref$ids$ensembl_mbl6_id)
nrow(ensembl_list_mcar)
length(mcar_replace_pos)

# how many of those do not match
table(is.na(mcar_insp_fill_pos))
table(is.na(mcar_replace_pos))
# different numbers of matches

# Look at the matches
# the number of matches is not the same from the different angles, there are
# 500 more matches in mcar_replace_pos than in mcar_insp_fill_pos
# these are likely duplicates

table(is.na(match(
  ensembl_list_mcar$mmusculus_homolog_ensembl_gene[!is.na(mcar_replace_pos)],
  insp_ref$ids$ensembl_mbl6_id[!is.na(mcar_insp_fill_pos)])))

table(is.na(match(
  insp_ref$ids$ensembl_mbl6_id[!is.na(mcar_insp_fill_pos)],
  ensembl_list_mcar$mmusculus_homolog_ensembl_gene[!is.na(mcar_replace_pos)])))
# non-NA IDs always match, so it is duplicated IDs that make the difference
# I will deal with duplicated IDs later, but keep them for reference

# Look at the NAs

# 28194 NAs
# it is most likely that these are simply IDs that are not annotated for mcar
# because the number of IDs in insp_genes is higher by the number of NAs 
# substracting 1 from ensembl_list_mcar because "" count as unique but not as id
nrow(insp_ref$ids) - length(unique(
  ensembl_list_mcar$mmusculus_homolog_ensembl_gene)[-1])

# they will stay empty or will later be referenced to other data
insp_ref$ids[which(is.na(mcar_insp_fill_pos)),]
insp_ref$gene_names[which(is.na(mcar_insp_fill_pos)),]

# 9822 NAs 
# this happens to be the same number as empty mmusculus_homolog_ensembl_gene
mcar_nas <- ensembl_list_mcar[which(is.na(mcar_replace_pos)),]

nrow(mcar_nas[mcar_nas$mmusculus_homolog_ensembl_gene != "",])
nrow(mcar_nas[mcar_nas$mmusculus_homolog_ensembl_gene == "",])

# indeed, the 9822 are empty
# I will deal with empty homolog_ensembl ids later

```

```{r add_mcar, message = FALSE}

# add the ensembl_mcar_id
insp_ref$ids$ensembl_mcar_id <- ensembl_list_mcar$ensembl_gene_id[
  mcar_insp_fill_pos]
insp_ref$ids

# in the same positions, add the homologous gene names
insp_ref$gene_names$mcar_mmusculus_homolog_associated_gene_name <- ensembl_list_mcar$mmusculus_homolog_associated_gene_name[mcar_insp_fill_pos]

# turn all "" in the mcar gene_names into NAs
unique(insp_ref$gene_names$mcar_mmusculus_homolog_associated_gene_name)[1:20]
insp_ref$gene_names$mcar_mmusculus_homolog_associated_gene_name[
  insp_ref$gene_names$mcar_mmusculus_homolog_associated_gene_name == ""] <- NA
insp_ref$gene_names

# also turn all "" in the mbl6_external_gene_name into NAs
insp_ref$gene_names$mbl6_external_gene_name[
  insp_ref$gene_names$mbl6_external_gene_name == ""] <- NA
insp_ref

```

### Castaneus

```{r add_mcas_check, message = FALSE}

head(ensembl_list_mcas)
nrow(ensembl_list_mcas)

# insp_ref positions that will be filled with corresponding data from list_mcas
mcas_insp_fill_pos <- match(insp_ref$ids$ensembl_mbl6_id, 
                            ensembl_list_mcas$mmusculus_homolog_ensembl_gene)
# should be same length is insp rows
nrow(insp_ref$ids)
length(mcas_insp_fill_pos)

# positions for ensembl_list_mcas that will replace empty slots in insp_ref
mcas_replace_pos <- match(ensembl_list_mcas$mmusculus_homolog_ensembl_gene,
                          insp_ref$ids$ensembl_mbl6_id)
nrow(ensembl_list_mcas)
length(mcas_replace_pos)

# how many of those do not match
table(is.na(mcas_insp_fill_pos))
table(is.na(mcas_replace_pos))
# the numbers are similar to mcar

# Look at the matches
table(is.na(match(
  ensembl_list_mcas$mmusculus_homolog_ensembl_gene[!is.na(mcas_replace_pos)],
  insp_ref$ids$ensembl_mbl6_id[!is.na(mcas_insp_fill_pos)])))

table(is.na(match(
  insp_ref$ids$ensembl_mbl6_id[!is.na(mcas_insp_fill_pos)],
  ensembl_list_mcas$mmusculus_homolog_ensembl_gene[!is.na(mcas_replace_pos)])))
# non-NA IDs always match, so it should be duplicated IDs that make the difference in numbers
# I will deal with duplicates later, but keep them for reference

# Look at the NAs

# 30038 NAs in mcas_insp_fill_pos
# some of these are simply IDs that are not annotated for mcas
# there are more NAs than the difference between insp_ref and mcas
# some of the IDs in mcas do not match to IDs in insp_ref
insp_ref$ids[is.na(mcas_insp_fill_pos),]
# I will deal with the non-matching mcas IDs later

# 13706 NAs in mcas_IDs
# most of these are empty
mcas_nas <- ensembl_list_mcas[which(is.na(mcas_replace_pos)),]

nrow(mcas_nas[mcas_nas$mmusculus_homolog_ensembl_gene != "",])
nrow(mcas_nas[mcas_nas$mmusculus_homolog_ensembl_gene == "",])

# 2359 of these NAs are not caused by empty gene names

# I will deal with the empty IDs later
# I will deal with the non-matching IDs later
# if they cannot be mapped to bl6 they probably won't be needed 

``` 

I have a strong suspicion that the 2359 IDs from ensembl_list_mcas that do not 
match the IDs insp_ref and some of the 30038 IDs IDs in insp_ref that do
not match ensembl_list_mcas actually describe the same genes but for some 
reason have different, non-matching IDs.

mcas have always posed problems.

I will proceed with the easy matching and get back to the mcas case later.

```{r add_mcas, message = FALSE}

# add the ensembl_mcas_id
insp_ref$ids$ensembl_mcas_id <- ensembl_list_mcas$ensembl_gene_id[
  mcas_insp_fill_pos]
insp_ref$ids

# in the same positions, add the homologous gene names
insp_ref$gene_names$mcas_mmusculus_homolog_associated_gene_name <- ensembl_list_mcas$mmusculus_homolog_associated_gene_name[mcas_insp_fill_pos]

# turn all "" in the mcar gene_names into NAs
unique(insp_ref$gene_names$mcas_mmusculus_homolog_associated_gene_name)[1:20]
insp_ref$gene_names$mcas_mmusculus_homolog_associated_gene_name[
  insp_ref$gene_names$mcas_mmusculus_homolog_associated_gene_name == ""] <- NA
insp_ref$gene_names

```

### Spretus

```{r add_mspr_check, message = FALSE}

head(ensembl_list_mspr)
nrow(ensembl_list_mspr)

# insp_ref positions that will be filled with corresponding data from list_mcar
mspr_insp_fill_pos <- match(insp_ref$ids$ensembl_mbl6_id, 
                            ensembl_list_mspr$mmusculus_homolog_ensembl_gene)
# should be same length is insp rows
nrow(insp_ref$ids)
length(mspr_insp_fill_pos)

# positions for ensembl_list_mspr that will replace empty slots in insp_ref
mspr_replace_pos <- match(ensembl_list_mspr$mmusculus_homolog_ensembl_gene,
                          insp_ref$ids$ensembl_mbl6_id)
nrow(ensembl_list_mspr)
length(mspr_replace_pos)

# how many of those do not match
table(is.na(mspr_insp_fill_pos))
table(is.na(mspr_replace_pos))

# Look at the matches
table(is.na(match(
  ensembl_list_mspr$mmusculus_homolog_ensembl_gene[!is.na(mspr_replace_pos)],
  insp_ref$ids$ensembl_mbl6_id[!is.na(mspr_insp_fill_pos)])))

table(is.na(match(
  insp_ref$ids$ensembl_mbl6_id[!is.na(mspr_insp_fill_pos)],
  ensembl_list_mspr$mmusculus_homolog_ensembl_gene[!is.na(mspr_replace_pos)])))
# non-NA IDs always match, so it is duplicated IDs that make the difference 
# I will deal with duplicated IDs later, but keep them for reference

# Look at the NAs

# 27419 NAs
# it is most likely that these are simply IDs that are not annotated for mspr
# because the number of IDs in insp_genes is higher by the number of NAs 
# substracting 1 from ensembl_list_mspr because "" counts as unique but not as id
nrow(insp_ref$ids) - length(unique(
  ensembl_list_mspr$mmusculus_homolog_ensembl_gene)[-1])

# they will stay empty or will later be referenced to other data
insp_ref$ids[which(is.na(mspr_insp_fill_pos)),]
insp_ref$gene_names[which(is.na(mspr_insp_fill_pos)),]


# 10665 NAs 
# this happens to be the same number as empty mmusculus_homolog_ensembl_gene
mspr_nas <- ensembl_list_mspr[which(is.na(mspr_replace_pos)),]

nrow(mspr_nas[mspr_nas$mmusculus_homolog_ensembl_gene != "",])
nrow(mspr_nas[mspr_nas$mmusculus_homolog_ensembl_gene == "",])

# indeed, the 10665 are empty
# I will deal with empty homolog_ensembl ids later

```

```{r add_mspr, message = FALSE}

# add the ensembl_mspr_id
insp_ref$ids$ensembl_mspr_id <- ensembl_list_mspr$ensembl_gene_id[
  mspr_insp_fill_pos]
insp_ref$ids

# in the same positions, add the homologous gene names
insp_ref$gene_names$mspr_mmusculus_homolog_associated_gene_name <- ensembl_list_mspr$mmusculus_homolog_associated_gene_name[mspr_insp_fill_pos]

# turn all "" in the mspr gene_names into NAs
unique(insp_ref$gene_names$mspr_mmusculus_homolog_associated_gene_name)[1:20]
insp_ref$gene_names$mspr_mmusculus_homolog_associated_gene_name[
  insp_ref$gene_names$mspr_mmusculus_homolog_associated_gene_name == ""] <- NA
insp_ref$gene_names

```

### Save Preliminary

```{r save_prelim, message = FALSE}

list.save(insp_ref, file = "../data/insp_ref_001.rds")

```

### Check data

```{r load_prelim, message = FALSE}

insp_ref <- list.load(file = "../data/insp_ref_001.rds")

```

```{r check_insp_ref, message = FALSE}

# Check the IDs

table(duplicated(insp_ref$ids$ensembl_mbl6_id))
table(duplicated(insp_ref$ids$ensembl_mcar_id))
table(duplicated(insp_ref$ids$ensembl_mcas_id))
table(duplicated(insp_ref$ids$ensembl_mspr_id))
# there are appoximately 25,000 unique IDs for each exotic species

table(insp_ref$ids$ensembl_mbl6_id == "") 
table(is.na(insp_ref$ids$ensembl_mbl6_id)) 
table(is.na(insp_ref$ids$ensembl_mcar_id)) 
table(is.na(insp_ref$ids$ensembl_mcas_id)) 
table(is.na(insp_ref$ids$ensembl_mspr_id)) 
# the numbers of IDs that duplicated and of those that are empty are similar

table(duplicated(insp_ref$ids$ensembl_mbl6_id[
  which(!is.na(insp_ref$ids$ensembl_mbl6_id))]))
table(duplicated(insp_ref$ids$ensembl_mcar_id[
  which(!is.na(insp_ref$ids$ensembl_mcar_id))]))
table(duplicated(insp_ref$ids$ensembl_mcas_id[
  which(!is.na(insp_ref$ids$ensembl_mcas_id))]))
table(duplicated(insp_ref$ids$ensembl_mspr_id[
  which(!is.na(insp_ref$ids$ensembl_mspr_id))]))
# from this mapping, there are approx. 500 duplicated IDs in the exotic species


# How many ensmus IDs have mapping to IDs from all three species?
filled_id_mcar <- which(!is.na(insp_ref$ids$ensembl_mcar_id))
filled_id_mcas <- which(!is.na(insp_ref$ids$ensembl_mcas_id))
filled_id_mspr <- which(!is.na(insp_ref$ids$ensembl_mspr_id))

filled_id_mcarmcas <- filled_id_mcar[which(!is.na(filled_id_mcar[
  match(filled_id_mcar, filled_id_mcas)]))]
filled_id_mcarmspr <- filled_id_mcar[which(!is.na(filled_id_mcar[
  match(filled_id_mcar, filled_id_mspr)]))]

filled_id_all <- filled_id_mcarmcas[which(!is.na(filled_id_mcarmcas[
  match(filled_id_mcarmcas, filled_id_mcarmspr)]))]

insp_ref$ids[filled_id_all,]
nrow(insp_ref$ids[filled_id_all,])
# approx. 20511 genes have shared IDs for all


# Check  the Gene names

table(duplicated(insp_ref$gene_names$mbl6_external_gene_name))
table(duplicated(
  insp_ref$gene_names$mcar_mmusculus_homolog_associated_gene_name))
table(duplicated(
  insp_ref$gene_names$mcas_mmusculus_homolog_associated_gene_name))
table(duplicated(
  insp_ref$gene_names$mspr_mmusculus_homolog_associated_gene_name))
# there are appoximately 25,000 unique IDs for each exotic species EXCEPT CAST

table(is.na(insp_ref$gene_names$mbl6_external_gene_name)) 
table(is.na(insp_ref$gene_names$mcar_mmusculus_homolog_associated_gene_name)) 
table(is.na(insp_ref$gene_names$mcas_mmusculus_homolog_associated_gene_name)) 
table(is.na(insp_ref$gene_names$mspr_mmusculus_homolog_associated_gene_name)) 
# there are 1492 gene names missing

table(duplicated(insp_ref$gene_names$mbl6_external_gene_name[which(!is.na(
  insp_ref$gene_names$mbl6_external_gene_name))]))
table(duplicated(insp_ref$gene_names$mcar_mmusculus_homolog_associated_gene_name[
  which(!is.na(
    insp_ref$gene_names$mcar_mmusculus_homolog_associated_gene_name))]))
table(duplicated(insp_ref$gene_names$mcas_mmusculus_homolog_associated_gene_name[
  which(!is.na(
    insp_ref$gene_names$mcas_mmusculus_homolog_associated_gene_name))]))
table(duplicated(insp_ref$gene_names$mspr_mmusculus_homolog_associated_gene_name[
  which(!is.na(
    insp_ref$gene_names$mspr_mmusculus_homolog_associated_gene_name))]))
# from this mapping, there are approx. 1500 duplicated gene names in the dataset


# are there any gene names in the other species where the bl6 names are NAs?

insp_ref_bl6nas <- insp_ref
insp_ref_bl6nas$gene_names <- insp_ref_bl6nas$gene_names[
  which(is.na(insp_ref$gene_names$mbl6_external_gene_name)),]

unique(insp_ref_bl6nas$gene_names$mcar_mmusculus_homolog_associated_gene_name)
unique(insp_ref_bl6nas$gene_names$mcas_mmusculus_homolog_associated_gene_name)
unique(insp_ref_bl6nas$gene_names$mspr_mmusculus_homolog_associated_gene_name)
# no, no other gene names could fill the mbl6 NAs

```
