---
author: "lea wölbert"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, message = FALSE}
library(SingleCellExperiment)
library(biomaRt)
library(rlist)
library(rtracklayer)
```

In this script I will deal with everything I promised I would deal with later
in 02_make_reftable.

```{r load_biomart, message = FALSE}

ensembl_mouse <- useMart(biomart="ENSEMBL_MART_ENSEMBL",
                         host="www.ensembl.org",
                         dataset="mmusculus_gene_ensembl")
ensembl_list_mbl6 <- getBM(attributes=c(
  "ensembl_gene_id",
  "external_gene_name"), 
  mart=ensembl_mouse)

ensembl_mcar <- useMart(biomart="ENSEMBL_MART_ENSEMBL",
                        host="www.ensembl.org",
                        dataset="mcaroli_gene_ensembl")
ensembl_list_mcar <- getBM(attributes=c(
  "ensembl_gene_id",
  "mmusculus_homolog_ensembl_gene",
  "mmusculus_homolog_associated_gene_name"), 
  mart=ensembl_mcar)

ensembl_mcas <- useMart(biomart="ENSEMBL_MART_MOUSE",
                        host="www.ensembl.org",
                        dataset="mmcasteij_gene_ensembl")
ensembl_list_mcas <- getBM(attributes=c(
  "ensembl_gene_id",
  "mmusculus_homolog_ensembl_gene",
  "mmusculus_homolog_associated_gene_name"), 
  mart=ensembl_mcas)

ensembl_mspr <- useMart(biomart="ENSEMBL_MART_ENSEMBL",
                        host="www.ensembl.org",
                        dataset="mspretus_gene_ensembl")
ensembl_list_mspr <- getBM(attributes=c(
  "ensembl_gene_id",
  "mmusculus_homolog_ensembl_gene",
  "mmusculus_homolog_associated_gene_name"),
  mart=ensembl_mspr)

```

```{r load_prelim, message = FALSE}

insp_ref <- list.load(file = "../../data/databases/insp_ref_001.rds")

```

```{r load_sce_objects, message = FALSE}

# objects merged (new) 
# THESE ARE DEPRECATED OBJECTS!!!! but the alignment is still the same for now
mbl6_old<- readRDS(file = "../../data/sce_objects/mbl6/mmus_old_00C")
mcar_old <- readRDS(file = "../../data/sce_objects/mcar/mcar_old_00C")
mcas_old <- readRDS(file = "../../data/sce_objects/mcas/mcas_old_00C")
mspr_old <- readRDS(file = "../../data/sce_objects/mspr/mspr_old_00C")

# objects merged (old)
mbl6_renorm <- readRDS(
  file = "../../data/sce_objects/mbl6/renorm_bl6_total_00C")
mcar_renorm <- readRDS(
  file = "../../data/sce_objects/mcar/renorm_caro_total_00C")
mcas_renorm <- readRDS(
  file = "../../data/sce_objects/mcas/renorm_cast_total_00C")
mspr_renorm <- readRDS(
  file = "../../data/sce_objects/mspr/renorm_spret_total_00C")

# other unmerged objects for reference 
mbl6_bm2_unm <- readRDS(file = "../../data/sce_objects/mbl6/bl61_bm2_006")
mbl6_bm3_unm <- readRDS(file = "../../data/sce_objects/mbl6/bl61_bm3_006")
mcas_bm2_unm <- readRDS(file = "../../data/sce_objects/mcas/cast1_bm2_006")
mcas_bm3_unm <- readRDS(file = "../../data/sce_objects/mcas/cast1_bm3_006")
mspr_bm2_unm <- readRDS(
  file = "../../data/sce_objects/mspr/pilot1_sce_mspr1_bm2")

```

```{r load_gtf_objects, message = FALSE}

# objects merged  both fractions
mcar_gtf <- rtracklayer::import(
  "../../data/databases/Mus_caroli.CAROLI_EIJ_v1.1.104.gtf")
mcar_gtf <- as.data.frame(mcar_gtf)
mcar_gtf <- mcar_gtf[,c(6, 10, 23)]
mcar_gtf <- mcar_gtf[duplicated(mcar_gtf$gene_id) != TRUE,]

mcas_gtf <- rtracklayer::import(
  "../../data/databases/Mus_musculus_casteij.CAST_EiJ_v1.104.gtf")
mcas_gtf <- as.data.frame(mcas_gtf)
mcas_gtf <- mcas_gtf[,c(6, 10, 12)]
mcas_gtf <- mcas_gtf[duplicated(mcas_gtf$gene_id) != TRUE,]


mspr_gtf <- rtracklayer::import(
  "../../data/databases/Mus_spretus.SPRET_EiJ_v1.104.gtf")
mspr_gtf <- as.data.frame(mspr_gtf)
mspr_gtf <- mspr_gtf[,c(6, 10, 23)]
mspr_gtf <- mspr_gtf[duplicated(mspr_gtf$gene_id) != TRUE,]

```

Previous main problems of matching data:

- there were many more insp_ref bl6 ensembl IDs than homolog ensembl IDs of 
exotic species

- there were exotic ensembl IDs that did not have a homolog ensembl ID to match 
to bl6 ensembl (these were "empty" IDs)

- For mcas, there were presumably homolog ensembl IDs in the mcas_ensembl_list 
that do not match any bl6 ensembl ID in insp_ref

- mcas have proven to be problematic in other aspects, too


I will first expand the insp_ref to add even the "empty" IDs for the sake of 
completeness.

Next, I will try to find out more about the non-matching IDs in Cast and decide
whether I should add those separately.
I will also check if these IDs are also found in the SCE objects.

Then, I will add info that is contained in the SCE objects.

Finally, I will add flags about duplicates into the metadata$duplicates slot to
warn about potential other IDs/sources/etc in the final database and give
the opportunity to check for those also.


# Expand rows using ensembl 

Although these genes in the exotic mice don't have an homolog ensembl ID,
they might still contain some useful information.

## Caroli

```{r make_empty_list_mcar, message = FALSE}

# the leftover mcar IDs are the ones that are empty in the mmus ID col
ensembl_list_mcar_empty <- ensembl_list_mcar
ensembl_list_mcar_empty <- ensembl_list_mcar_empty[
  ensembl_list_mcar_empty$mmusculus_homolog_ensembl_gene == "",]
nrow(insp_ref$ids)
nrow(ensembl_list_mcar_empty)

unique(ensembl_list_mcar_empty$mmusculus_homolog_associated_gene_name)
# no gene names contained, so only mcar ensmus IDs would be added

# first check if the leftover mcar ensmus IDs are already part of the insp_ref
table(is.na(match(ensembl_list_mcar_empty$ensembl_gene_id, 
                   insp_ref$ids$ensembl_mcar_id)))
# no

```

```{r add_empty_list_mcar}

# then just rbind an object of same structure as insp_ref 
# but containing only these rows to the actual insp_ref

l_mcar <- nrow(ensembl_list_mcar_empty)

insp_ref_add_mcar <- list(
  ids = data.frame(
    ensembl_mbl6_id = vector("character", length = l_mcar),
    ensembl_mcar_id = ensembl_list_mcar_empty$ensembl_gene_id,
    ensembl_mcas_id = vector("character", length =l_mcar),
    ensembl_mspr_id = vector("character", length = l_mcar)
   ),
  gene_names = data.frame(
    ensembl_mbl6_id = vector("character", length = l_mcar),
    mbl6_external_gene_name = vector("character", length = l_mcar),
    mcar_mmusculus_homolog_associated_gene_name = vector("character", 
                                                         length = l_mcar),
    mcas_mmusculus_homolog_associated_gene_name = vector("character", length = 
                                                           l_mcar),
    mspr_mmusculus_homolog_associated_gene_name = vector("character", 
                                                         length = l_mcar)
    ),
   synonyms = data.frame(
     ensembl_mbl6_id = vector("character", length = l_mcar),
     mbl6_external_gene_name = vector("character", length = l_mcar),
     synonym.1 = vector("character", length = l_mcar),
     synonym.2 = vector("character", length = l_mcar),
     synonym.3 = vector("character", length = l_mcar),
     synonym.4 = vector("character", length = l_mcar),
     synonym.5 = vector("character", length = l_mcar),
     synonym.6 = vector("character", length = l_mcar)
    ),
    metadata = data.frame(
      ensembl_mbl6_id = vector("character", length = l_mcar),
      duplicated = vector("character", length =l_mcar),
      source = rep("ensembl_mcar", length = l_mcar),
      notes = vector("character", length = l_mcar)
    )
)

nrow(insp_ref_add_mcar$ids)

```

```{r rbind_empty_list_mcar, message = FALSE}

nrow(insp_ref$ids)
insp_ref$ids <- rbind(insp_ref$ids, insp_ref_add_mcar$ids)
insp_ref$gene_names  <- rbind(insp_ref$gene_names, insp_ref_add_mcar$gene_names)
insp_ref$synonyms <- rbind(insp_ref$synonyms, insp_ref_add_mcar$synonyms)
insp_ref$metadata <- rbind(insp_ref$metadata, insp_ref_add_mcar$metadata)

nrow(insp_ref$ids)
tail(insp_ref$metadata)
tail(insp_ref$ids)

```

## Castaneus

So mcas has some further problems outside of empty homolog bl6 ids.
However, there still are quite a few empty ones, so I will also add those

```{r make_empty_list_mcas, message = FALSE}

ensembl_list_mcas_empty <- ensembl_list_mcas
ensembl_list_mcas_empty <- ensembl_list_mcas_empty[
  ensembl_list_mcas_empty$mmusculus_homolog_ensembl_gene == "",]
nrow(insp_ref$ids)
nrow(ensembl_list_mcas_empty)

unique(ensembl_list_mcas_empty$mmusculus_homolog_associated_gene_name)
# no gene names contained, so only mcas ensmus IDs would be added

# first check if the leftover mcas ensmus IDs are already part of the insp_ref
table(!is.na(match(ensembl_list_mcas_empty$ensembl_gene_id, 
                   insp_ref$ids$ensembl_mcas_id)))
# no

```

```{r add_empty_list_mcas, message = FALSE}

# then just rbind a object of same structure as insp_ref 

l_mcas <- nrow(ensembl_list_mcas_empty)

insp_ref_add_mcas <- list(
  ids = data.frame(
    ensembl_mbl6_id = vector("character", length = l_mcas),
    ensembl_mcar_id = vector("character", length = l_mcas),
    ensembl_mcas_id = ensembl_list_mcas_empty$ensembl_gene_id,
    ensembl_mspr_id = vector("character", length = l_mcas)
   ),
  gene_names = data.frame(
    ensembl_mbl6_id = vector("character", length = l_mcas),
    mbl6_external_gene_name = vector("character", length = l_mcas),
    mcar_mmusculus_homolog_associated_gene_name = vector("character", 
                                                         length = l_mcas),
    mcas_mmusculus_homolog_associated_gene_name = vector("character", 
                                                         length = l_mcas),
    mspr_mmusculus_homolog_associated_gene_name = vector("character", 
                                                         length = l_mcas)
    ),
  synonyms = data.frame(
    ensembl_mbl6_id = vector("character", length = l_mcas),
    mbl6_external_gene_name = vector("character", length = l_mcas),
    synonym.1 = vector("character", length = l_mcas),
    synonym.2 = vector("character", length = l_mcas),
    synonym.3 = vector("character", length = l_mcas),
    synonym.4 = vector("character", length = l_mcas),
    synonym.5 = vector("character", length = l_mcas),
    synonym.6 = vector("character", length = l_mcas)
    ),
  metadata = data.frame(ensembl_mbl6_id = vector("character", length = l_mcas),
    duplicated = vector("character", length = l_mcas),
    source = rep("ensembl_mcas", length = l_mcas),
    notes = vector("character", length = l_mcas)
   )
 )

nrow(insp_ref_add_mcas$ids)

```

```{r rbind_empty_list_mcas, message = FALSE}

nrow(insp_ref$ids)
insp_ref$ids <- rbind(insp_ref$ids, insp_ref_add_mcas$ids)
insp_ref$gene_names  <- rbind(insp_ref$gene_names, insp_ref_add_mcas$gene_names)
insp_ref$synonyms <- rbind(insp_ref$synonyms, insp_ref_add_mcas$synonyms)
insp_ref$metadata <- rbind(insp_ref$metadata, insp_ref_add_mcas$metadata)

nrow(insp_ref$ids)
tail(insp_ref$metadata)
tail(insp_ref$ids)

```

## Spretus

```{r make_empty_list_mspr, message = FALSE}

ensembl_list_mspr_empty <- ensembl_list_mspr
ensembl_list_mspr_empty <- ensembl_list_mspr_empty[
  ensembl_list_mspr_empty$mmusculus_homolog_ensembl_gene == "",]
nrow(insp_ref$ids)
nrow(ensembl_list_mspr_empty)

unique(ensembl_list_mspr_empty$mmusculus_homolog_associated_gene_name)
# no gene names contained, so only mspr ensmus IDs would be added

# first check if the leftover mspr ensmus IDs are already part of the insp_ref
table(!is.na(match(ensembl_list_mspr_empty$ensembl_gene_id, 
                   insp_ref$ids$ensembl_mspr_id)))
# no

```

```{r add_empty_list_mspr, message = FALSE}

l_mspr <- nrow(ensembl_list_mspr_empty)

insp_ref_add_mspr <- list(
  ids = data.frame(
    ensembl_mbl6_id = vector("character", length = l_mspr),
    ensembl_mcar_id = vector("character", length = l_mspr),
    ensembl_mcas_id = vector("character", length = l_mspr),
    ensembl_mspr_id = ensembl_list_mspr_empty$ensembl_gene_id
   ),
  gene_names = data.frame(
    ensembl_mbl6_id = vector("character", length = l_mspr),
    mbl6_external_gene_name = vector("character", length = l_mspr),
    mcar_mmusculus_homolog_associated_gene_name = vector("character",
                                                         length = l_mspr),
    mcas_mmusculus_homolog_associated_gene_name = vector("character", 
                                                         length = l_mspr),
    mspr_mmusculus_homolog_associated_gene_name = vector("character",
                                                         length = l_mspr)
    ),
  synonyms = data.frame(
    ensembl_mbl6_id = vector("character", length = l_mspr),
    mbl6_external_gene_name = vector("character", length = l_mspr),
    synonym.1 = vector("character", length = l_mspr),
    synonym.2 = vector("character", length = l_mspr),
    synonym.3 = vector("character", length = l_mspr),
    synonym.4 = vector("character", length = l_mspr),
    synonym.5 = vector("character", length = l_mspr),
    synonym.6 = vector("character", length = l_mspr)
   ),
  metadata = data.frame(
    ensembl_mbl6_id = vector("character", length = l_mspr),
    duplicated = vector("character", length = l_mspr),
    source = rep("ensembl_mspr", length = l_mspr),
    notes = vector("character", length = l_mspr)
   )
)

nrow(insp_ref_add_mspr$ids)

```

```{r rbind_empty_list_mspr, message = FALSE}

nrow(insp_ref$ids)
insp_ref$ids <- rbind(insp_ref$ids, insp_ref_add_mspr$ids)
insp_ref$gene_names  <- rbind(insp_ref$gene_names, insp_ref_add_mspr$gene_names)
insp_ref$synonyms <- rbind(insp_ref$synonyms, insp_ref_add_mspr$synonyms)
insp_ref$metadata <- rbind(insp_ref$metadata, insp_ref_add_mspr$metadata)

nrow(insp_ref$ids)
tail(insp_ref$metadata)
tail(insp_ref$ids)

```

## Checkpoint 

```{r save_cp_1, message = FALSE}

list.save(insp_ref, file = "../../data/databases/insp_ref_002.rds")

```


```{r load_cp_1, message = FALSE}

insp_ref <- list.load(file = "../../data/databases/insp_ref_002.rds")

```

# Deal with Cast non-identical IDs

There were some non-identical IDs between ensembl_list_mcas and insp_ref
that could not be explained by empty IDs or higher number of total IDs in 
insp_ref.

## Should I add?

First, I isolate the non-matchable IDs from ensembl_list_mcas

```{r make_matches, message = FALSE}

mcas_insp_fill_pos <- match(insp_ref$ids$ensembl_mbl6_id, 
                            ensembl_list_mcas$mmusculus_homolog_ensembl_gene)
# should be same length as insp rows
nrow(insp_ref$ids)
length(mcas_insp_fill_pos)

# positions for ensembl_list_mcas that will replace empty slots in insp_ref
mcas_replace_pos <- match(ensembl_list_mcas$mmusculus_homolog_ensembl_gene,
                          insp_ref$ids$ensembl_mbl6_id)
nrow(ensembl_list_mcas)
length(mcas_replace_pos)

# how many of those do not match
table(is.na(mcas_insp_fill_pos))
table(is.na(mcas_replace_pos))
# around 2400 IDs in are in ensembl_list_mcas but not in insp_ref
# even after the ones without mmus ensmus IDs were added by rbind

```

First I look at the IDs in ensembl_list_mcas that cannot be matched to insp_ref

```{r find_mcas_nas_subset, message = FALSE}

# subset ensembl_list_mcas to only contain these NA IDs

ensembl_list_mcas_nas <- ensembl_list_mcas
ensembl_list_mcas_nas <-  ensembl_list_mcas_nas[which(is.na(mcas_replace_pos)),]
nrow(ensembl_list_mcas_nas)
head(ensembl_list_mcas_nas)
# they contain homolog ensembl gene ids that could NOT be matched to insp_ref

```

```{r mcas_nas_gene_names, message = FALSE}

# Look at the gene names
length(unique(ensembl_list_mcas_nas$mmusculus_homolog_associated_gene_name))

# extract the unique gene names from the non-matching IDs that are not ""
gene_names_mcas_na_ids <- ensembl_list_mcas_nas$mmusculus_homolog_associated_gene_name
gene_names_mcas_na_ids <- unique(gene_names_mcas_na_ids)
gene_names_mcas_na_ids <- gene_names_mcas_na_ids[gene_names_mcas_na_ids != ""]
head(gene_names_mcas_na_ids)
# 426 gene names belong to the IDs that are not found in insp_ref

# some of mcas gene names were highly duplicated, so some may already be in 
# insp_ref after adding them with the "easy" or "empty" matches from before


# Look at gene names that are not exclusive to ensembl_list_mcas

# number of gene names in ensembl_list_mcas_nas already contained in insp_ref
table(!is.na(match(
  gene_names_mcas_na_ids, insp_ref$gene_names$mbl6_external_gene_name)))
table(!is.na(match(
  gene_names_mcas_na_ids, 
  insp_ref$gene_names$mcas_mmusculus_homolog_associated_gene_name)))
# 40 of those gene names are already in mbl6_external_gene_name
# and 69 are also in mcas_mmusculus_homolog_associated_gene_name
# they are probably alternative gene names/synonyms

```

```{r mcas_nas_gn_check, message = FALSE}

# look at potential alternative names

# subset insp_ref to show rows with these gene names 
head(insp_ref$gene_names[!is.na(match(
  insp_ref$gene_names$mcas_mmusculus_homolog_associated_gene_name, 
  gene_names_mcas_na_ids)),])
shared_gene_names1 <- insp_ref$gene_names$mcas_mmusculus_homolog_associated_gene_name[
  !is.na(match(insp_ref$gene_names$mcas_mmusculus_homolog_associated_gene_name, 
               gene_names_mcas_na_ids))]
shared_gene_names1 <- unique(shared_gene_names1)
head(shared_gene_names1)
# many of these rows have Gms (predicted genes) listed as ext gene name
# but in the mcas reference, they have actual names

# subset insp_ref to show rows with these gene names in mbl6_external_gene_name
shared_gene_names2 <- insp_ref$gene_names$mbl6_external_gene_name[!is.na(match(
  insp_ref$gene_names$mbl6_external_gene_name, gene_names_mcas_na_ids))]
shared_gene_names2 <- unique(shared_gene_names2)
head(shared_gene_names2)
# most of those are some rRNAs, snoRNAs, small nuclear RNAs

# match to get the gene names that are in both 
# mcas_mmusculus_homolog_associated_gene_name and mbl6_external_gene_name
# interestingly, there are no genes that are in both
shared_gene_names <- shared_gene_names1[which(!is.na(
  match(shared_gene_names1, shared_gene_names2)))]

```

```{r mcas_nas_gn_check_sce, message = FALSE}

# check if these gene names are found in the sce objects
length(which(!is.na(match(gene_names_mcas_na_ids, 
                          rownames(mcas_old)))))
# of course found in cast

length(which(!is.na(match(gene_names_mcas_na_ids, 
                          rownames(mbl6_old)))))
# also found in bl6

length(which(!is.na(match(gene_names_mcas_na_ids, 
                          rownames(mcar_old)))))
# not found in caroli

length(which(!is.na(match(gene_names_mcas_na_ids, 
                          rownames(mspr_old)))))
# one found in spret

```

```{r mcas_nas_gene_names_exclusive, message = FALSE}

# Look at the gene names that are still exclusive to ensembl_list_mcas_nas
# = gene names that are not yet found in insp_ref

excls_gene_names1 <- gene_names_mcas_na_ids[is.na(match(
  gene_names_mcas_na_ids, insp_ref$gene_names$mbl6_external_gene_name))]
excls_gene_names_uni1 <- unique(excls_gene_names1)
length(excls_gene_names_uni1)

excls_gene_names2 <- gene_names_mcas_na_ids[is.na(
  match(gene_names_mcas_na_ids, 
        insp_ref$gene_names$mcas_mmusculus_homolog_associated_gene_name))]
excls_gene_names_uni2 <- unique(excls_gene_names2)
length(excls_gene_names_uni2)

excls_gene_names <- excls_gene_names_uni1[which(!is.na(match(
  excls_gene_names_uni1, excls_gene_names_uni2)))]
length(excls_gene_names)
# these genes are found neither in insp_ref mbl6_external_gene_name 
# nor mcas_mmusculus_homolog_associated_gene_name

# check if these genes are found in any of the SCE objects
length(which(!is.na(match(excls_gene_names, rownames(mcas_old)))))
# of course found in cast

length(which(!is.na(match(excls_gene_names, rownames(mbl6_old)))))
# also found in bl6

length(which(!is.na(match(excls_gene_names, rownames(mcar_old)))))
length(which(!is.na(match(excls_gene_names, rownames(mspr_old)))))
# not found in the other exotic mice

```

Because these gene names are found in other objects, all additional IDs
from ensembl_list_mcas will be added to insp_ref.

## Add them

```{r mcas_nas_ids, message = FALSE}

# Look at the gene names
length(unique(ensembl_list_mcas_nas$ensembl_gene_id))

# extract the unique gene names from the non-matching IDs that are not ""
ids_mcas_na <- ensembl_list_mcas_nas$ensembl_gene_id
ids_mcas_na <- unique(ids_mcas_na)
ids_mcas_na <- ids_mcas_na[ids_mcas_na != ""]
length(ids_mcas_na)
# 2340 gene names belong to the IDs that are not found in insp_ref

# some of mcas gene names were highly duplicated, so some may already be in 
# insp_ref after adding them with the "easy" or "empty" matches from before

```

```{r mcas_nas_ids_check, message = FALSE}

# which nas IDs are found in both datasets?
# those were likely duplicated in ensembl_list_mcas
shared_ids_mcas_na <- ids_mcas_na[which(!is.na(match(
  ids_mcas_na, insp_ref$ids$ensembl_mcas_id )))]
length(shared_ids_mcas_na)
# only four of those were previously duplicated

insp_ref$ids[which(!is.na(match(insp_ref$ids$ensembl_mcas_id, 
                                shared_ids_mcas_na))),]
insp_ref$gene_names[which(!is.na(match(insp_ref$ids$ensembl_mcas_id, 
                                       shared_ids_mcas_na))),]
ensembl_list_mcas_nas[which(!is.na(match(ensembl_list_mcas_nas$ensembl_gene_id,
                                         shared_ids_mcas_na))),]

# they seem to correspond to the same genes in both datasets
excl_ids_mcas_na <- ids_mcas_na[which(is.na(match(
  ids_mcas_na,insp_ref$ids$ensembl_mcas_id)))]
excl_ids_mcas_na <- unique(excl_ids_mcas_na)
excl_ids_mcas_na <- excl_ids_mcas_na[excl_ids_mcas_na != ""]

# these are the cast IDs not found in insp_ref at all
length(excl_ids_mcas_na)

```

```{r mcas_nas_ids_check_sce, message = FALSE}

# do the rows contain more information?
head(ensembl_list_mcas_nas[which(!is.na(match(
  ensembl_list_mcas_nas$ensembl_gene_id,  excl_ids_mcas_na))),])
# they mostly contain the gene names which we already found

#check if these IDs are found in the SCE objects

length(which(!is.na(match(excl_ids_mcas_na, rownames(mcas_old)))))
# around a 100 of these found in cast

length(which(!is.na(match(excl_ids_mcas_na, rownames(mbl6_old)))))
length(which(!is.na(match(excl_ids_mcas_na, rownames(mcar_old)))))
length(which(!is.na(match(excl_ids_mcas_na, rownames(mspr_old)))))
# of course not found in the other mice

```

```{r add_cast_na_ids, message = FALSE}

l_mcas_nas <- nrow(ensembl_list_mcas_nas)

insp_ref_add_mcas_nas <- list(
  ids = data.frame(
    ensembl_mbl6_id = ensembl_list_mcas_nas$mmusculus_homolog_ensembl_gene,
    ensembl_mcar_id = vector("character", length = l_mcas_nas),
    ensembl_mcas_id = ensembl_list_mcas_nas$ensembl_gene_id,
    ensembl_mspr_id = vector("character", length = l_mcas_nas)
   ),
  gene_names = data.frame(
    ensembl_mbl6_id = ensembl_list_mcas_nas$mmusculus_homolog_ensembl_gene,
    mbl6_external_gene_name = vector("character", length = l_mcas_nas),
    mcar_mmusculus_homolog_associated_gene_name = vector("character", 
                                                         length = l_mcas_nas),
    mcas_mmusculus_homolog_associated_gene_name = ensembl_list_mcas_nas$mmusculus_homolog_associated_gene_name,
    mspr_mmusculus_homolog_associated_gene_name = vector("character", 
                                                         length = l_mcas_nas)
   ),
  synonyms = data.frame(
    ensembl_mbl6_id = ensembl_list_mcas_nas$mmusculus_homolog_ensembl_gene,
    mbl6_external_gene_name = vector("character", length = l_mcas_nas),
    synonym.1 = vector("character", length = l_mcas_nas),
    synonym.2 = vector("character", length = l_mcas_nas),
    synonym.3 = vector("character", length = l_mcas_nas),
    synonym.4 = vector("character", length = l_mcas_nas),
    synonym.5 = vector("character", length = l_mcas_nas),
    synonym.6 = vector("character", length = l_mcas_nas)
  ),
  metadata = data.frame(
    ensembl_mbl6_id = ensembl_list_mcas_nas$mmusculus_homolog_ensembl_gene,
    duplicated = vector("character", length = l_mcas_nas),
    source = rep("ensembl_mcas", length = l_mcas_nas),
    notes = rep("Possibly_deprecated")
   )
 )

nrow(insp_ref_add_mcas_nas$ids)

# INTERESTINGLY some ensmus IDs are duplicated WHEN DERIVED FROM CAST 
table(duplicated(insp_ref_add_mcas_nas$ids$ensembl_mbl6_id))

```

```{r rbind_cast_na_ids, message = FALSE}

nrow(insp_ref$ids)
insp_ref$ids <- rbind(insp_ref$ids, insp_ref_add_mcas_nas$ids)
insp_ref$gene_names  <- rbind(insp_ref$gene_names, 
                              insp_ref_add_mcas_nas$gene_names)
insp_ref$synonyms <- rbind(insp_ref$synonyms, insp_ref_add_mcas_nas$synonyms)
insp_ref$metadata <- rbind(insp_ref$metadata, insp_ref_add_mcas_nas$metadata)

nrow(insp_ref$ids)
tail(insp_ref$metadata)
tail(insp_ref$ids)

```

## Checkpoint

```{r save_cp_2, message = FALSE}

list.save(insp_ref, file = "../../data/databases/insp_ref_003.rds")

```

```{r load_cp_2, message = FALSE}

insp_ref <- list.load(file = "../../data/databases/insp_ref_003.rds")

```

# Add Info From SCE Objects

Is the info from all SCE objects identical?

```{r compare_between_sce_objects, message = FALSE}

# BL6
# compare between fractions of same species
table(is.na(match(rowData(mbl6_bm2_unm)$ID, rowData(mbl6_bm3_unm)$ID)))
table(is.na(match(rowData(mbl6_bm2_unm)$Symbol, rowData(mbl6_bm3_unm)$Symbol)))

# CAST 
table(is.na(match(rowData(mcas_bm2_unm)$ID, rowData(mcas_bm3_unm)$ID)))
table(is.na(match(rowData(mcas_bm2_unm)$Symbol, rowData(mcas_bm3_unm)$Symbol)))
# no difference between the fractions apparently

```

```{r compare_between_bl6, message = FALSE}

# BL6
# compare between merged and unmerged
table(is.na(match(rowData(mbl6_bm2_unm)$ID, rowData(mbl6_old)$ID)))
table(is.na(match(rowData(mbl6_old)$ID, rowData(mbl6_bm2_unm)$ID)))

table(is.na(match(rowData(mbl6_renorm)$ID, rowData(mbl6_old)$ID)))
table(is.na(match(rowData(mbl6_old)$ID, rowData(mbl6_renorm)$ID)))

table(is.na(match(rowData(mbl6_renorm)$ID, rowData(mbl6_bm2_unm)$ID)))
table(is.na(match(rowData(mbl6_bm2_unm)$ID, rowData(mbl6_renorm)$ID)))
# there are more rows in mbl6_old 
# these contain additional IDs compared to other objects
# mbl6_renorm and mbl6_bm2_unm have the same IDs

```

```{r compare_between_caro, message = FALSE}

# CARO
table(is.na(match(rowData(mcar_renorm)$ID, rowData(mcar_old)$ID)))
table(is.na(match(rowData(mcar_old)$ID, rowData(mcar_renorm)$ID)))
# no differences for caro

```

```{r compare_between_cast, message = FALSE}

# CAST
table(is.na(match(rowData(mcas_bm2_unm)$ID, rowData(mcas_old)$ID)))
table(is.na(match(rowData(mcas_old)$ID, rowData(mcas_bm2_unm)$ID)))

table(is.na(match(rowData(mcas_renorm)$ID, rowData(mcas_old)$ID)))
table(is.na(match(rowData(mcas_old)$ID, rowData(mcas_renorm)$ID)))

table(is.na(match(rowData(mcas_renorm)$ID, rowData(mcas_bm2_unm)$ID)))
table(is.na(match(rowData(mcas_bm2_unm)$ID, rowData(mcas_renorm)$ID)))
# there are additional unique IDs in mcas_bm2_unm and mcas_old
# mcas_renorm and mcas_old have the same IDs

```

```{r compare_between_spret, message = FALSE}

# SPRET
# IDs
table(is.na(match(rowData(mspr_bm2_unm)$ID, rowData(mspr_old)$ID)))
table(is.na(match(rowData(mspr_old)$ID, rowData(mspr_bm2_unm)$ID)))

table(is.na(match(rowData(mspr_renorm)$ID, rowData(mspr_old)$ID)))
table(is.na(match(rowData(mspr_old)$ID, rowData(mspr_renorm)$ID)))

table(is.na(match(rowData(mspr_renorm)$ID, rowData(mspr_bm2_unm)$ID)))
table(is.na(match(rowData(mspr_bm2_unm)$ID, rowData(mspr_renorm)$ID)))

# Between Symbol
table(is.na(match(rowData(mspr_bm2_unm)$Symbol, rowData(mspr_old)$Symbol)))
table(is.na(match(rowData(mspr_old)$Symbol, rowData(mspr_bm2_unm)$Symbol)))

table(is.na(match(rowData(mspr_renorm)$Symbol, rowData(mspr_old)$Symbol)))
table(is.na(match(rowData(mspr_old)$Symbol, rowData(mspr_renorm)$Symbol)))

table(is.na(match(rowData(mspr_renorm)$Symbol, rowData(mspr_bm2_unm)$Symbol)))
table(is.na(match(rowData(mspr_bm2_unm)$Symbol, rowData(mspr_renorm)$Symbol)))

# Between Symbols2 
table(is.na(match(rowData(mspr_bm2_unm)$Symbols2, rowData(mspr_old)$Symbols2)))
table(is.na(match(rowData(mspr_old)$Symbols2, rowData(mspr_bm2_unm)$Symbols2)))

table(is.na(match(rowData(mspr_renorm)$Symbols2, rowData(mspr_old)$Symbols2)))
table(is.na(match(rowData(mspr_old)$Symbols2, rowData(mspr_renorm)$Symbols2)))

table(is.na(match(rowData(mspr_renorm)$Symbols2, 
                  rowData(mspr_bm2_unm)$Symbols2)))
table(is.na(match(rowData(mspr_bm2_unm)$Symbols2, 
                  rowData(mspr_renorm)$Symbols2)))
# again no difference, how about between Symbol and Symbols2

# Compare Symbol to Symbols2
table(is.na(match(rowData(mspr_bm2_unm)$Symbol, 
                  rowData(mspr_bm2_unm)$Symbols2)))
table(is.na(match(rowData(mspr_bm2_unm)$Symbol, 
                  rowData(mspr_bm2_unm)$Symbols2)))

table(is.na(match(rowData(mspr_old)$Symbol, rowData(mspr_old)$Symbols2)))
table(is.na(match(rowData(mspr_old)$Symbol, rowData(mspr_old)$Symbols2)))

table(is.na(match(rowData(mspr_renorm)$Symbol, rowData(mspr_renorm)$Symbols2)))
table(is.na(match(rowData(mspr_renorm)$Symbol, rowData(mspr_renorm)$Symbols2)))
# A LOT OF DIFFERENCES
# The most important differences are between Symbol and Symbols2 
# of the same object

```

## BL6

```{r compare_bl6_sce, message = FALSE}

# three bl6 objects: mbl6_bm2_unm, mbl6_old, mbl6_renorm

#ids: is there more info in the SCE object
table(is.na(match(rowData(mbl6_bm2_unm)$ID, insp_ref$ids$ensembl_mbl6_id)))
table(is.na(match(rowData(mbl6_old)$ID, insp_ref$ids$ensembl_mbl6_id)))
table(is.na(match(rowData(mbl6_renorm)$ID, insp_ref$ids$ensembl_mbl6_id)))
# the overlap is ok (1000 missing IDs out of 50,000) for IDs

# gene names
table(is.na(match(rowData(mbl6_bm2_unm)$Symbol, 
                  insp_ref$gene_names$mbl6_external_gene_name)))
table(is.na(match(rowData(mbl6_old)$Symbol, 
                  insp_ref$gene_names$mbl6_external_gene_name)))
table(is.na(match(rowData(mbl6_renorm)$Symbol, 
                  insp_ref$gene_names$mbl6_external_gene_name)))
# there are fewer matching gene names
# around 5000 gene names are missing  in insp_ref

# mbl6_old has the highest number of missing IDs and names

```

### IDs

```{r compare_bl6_sce_ids, message = FALSE}

# look the the non-matching IDs
# there are around 1100 IDs in the SCE object which are not in the insp_ref
# we have established that mbl6_bm2_unm and mbl6_renorm match 
nm_ids_mbl6.1 <- rowData(mbl6_bm2_unm)$ID[which(is.na(match(
  rowData(mbl6_bm2_unm)$ID, insp_ref$ids$ensembl_mbl6_id)))]
length(nm_ids_mbl6.1)

nm_ids_mbl6.2 <- rowData(mbl6_old)$ID[which(is.na(match(
  rowData(mbl6_old)$ID, insp_ref$ids$ensembl_mbl6_id)))]
length(nm_ids_mbl6.2)

table(is.na(match(nm_ids_mbl6.2, nm_ids_mbl6.1)))
table(is.na(match(nm_ids_mbl6.1, nm_ids_mbl6.2)))
# there are 16 more IDs in nm_ids_mbl6.2, so use that one
# the rest of the IDs match

# look at the additional 16 IDs in nm_ids_mbl6.2
addtl_ids <- nm_ids_mbl6.2[which(is.na(match(nm_ids_mbl6.2, nm_ids_mbl6.1)))]
rowData(mbl6_old)[which(!is.na(match(rowData(mbl6_old)$ID,
                                        addtl_ids))),]

```

```{r add_bl6_sce_ids, message = FALSE}

nr_bl6 <- length(nm_ids_mbl6.2)
pos_bl6 <- which(!is.na(match(rowData(mbl6_old)$ID, nm_ids_mbl6.2)))

# none of the symbols to be added are actually IDs
length(grep("ENSMUS", rowData(mbl6_old)$Symbol))

# simply add these IDs using rbind
# the gene names are put gene names 
add_mbl6_sce <- list(
  ids = data.frame(
    ensembl_mbl6_id = nm_ids_mbl6.2,
    ensembl_mcar_id = vector("character", length = nr_bl6),
    ensembl_mcas_id = vector("character", length = nr_bl6),
    ensembl_mspr_id = vector("character", length = nr_bl6)
  ),
  gene_names = data.frame(
    ensembl_mbl6_id = nm_ids_mbl6.2,
    mbl6_external_gene_name = rowData(mbl6_old)$Symbol[pos_bl6],
    mcar_mmusculus_homolog_associated_gene_name = vector("character",
                                                         length = nr_bl6),
    mcas_mmusculus_homolog_associated_gene_name = vector("character", 
                                                         length = nr_bl6),
    mspr_mmusculus_homolog_associated_gene_name = vector("character", 
                                                         length = nr_bl6)
 ),
 synonyms = data.frame(
   ensembl_mbl6_id = nm_ids_mbl6.2,
   mbl6_external_gene_name = rowData(mbl6_old)$ID[pos_bl6],
   synonym.1 =  vector("character", length = nr_bl6),
   synonym.2 = vector("character", length = nr_bl6),
   synonym.3 = vector("character", length = nr_bl6),
   synonym.4 = vector("character", length = nr_bl6),
   synonym.5 = vector("character", length = nr_bl6),
   synonym.6 = vector("character", length = nr_bl6)
  ),
  metadata = data.frame(
    ensembl_mbl6_id = nm_ids_mbl6.2,
    duplicated = vector("character", length = nr_bl6),
    source = rep("BL6_SCE_object", length = nr_bl6),
    notes = rep("Possibly_deprecated", length =nr_bl6)
  )
)

head(add_mbl6_sce$ids)
head(add_mbl6_sce$gene_names)
nrow(add_mbl6_sce$ids)

# I added "possibly deprecated" because I had a hard time finding
# some of these IDs on the mgi website

```

```{r rbind_bl6_sce_ids, message = FALSE}

insp_ref$ids <- rbind(insp_ref$ids, add_mbl6_sce$ids)
insp_ref$gene_names  <- rbind(insp_ref$gene_names, add_mbl6_sce$gene_names)
insp_ref$synonyms <- rbind(insp_ref$synonyms, add_mbl6_sce$synonyms)
insp_ref$metadata <- rbind(insp_ref$metadata, add_mbl6_sce$metadata)

tail(insp_ref$ids)
tail(insp_ref$metadata)

```

### Gene Names

Add gene names as synonyms.

```{r compare_bl6_sce_names, message = FALSE}

# these gene names are only in BL6_SCE_object and not in insp_ref:
nm_gn_mbl6.1<- rowData(mbl6_bm2_unm)$Symbol[
  which(is.na(match(rowData(mbl6_bm2_unm)$Symbol, 
                    insp_ref$gene_names$mbl6_external_gene_name)))]
length(nm_gn_mbl6.1)

nm_gn_mbl6.2 <- rowData(mbl6_old)$Symbol[
  which(is.na(match(rowData(mbl6_old)$Symbol,
                    insp_ref$gene_names$mbl6_external_gene_name)))]
length(nm_gn_mbl6.2)
# around 4000 gene names are found in BL6_Sce_objects
# there are 6 more gene names in nm_gn_mbl6.2 than in nm_gn_mbl6.1

table(is.na(match(nm_gn_mbl6.1, nm_gn_mbl6.2)))
table(is.na(match(nm_gn_mbl6.2, nm_gn_mbl6.1)))
# but all gene names seem to match
table(duplicated(nm_gn_mbl6.2))
table(duplicated(rownames(mbl6_old)))
# there are some duplicates in the mbl6_old symbols and rownames
# use nm_gn_mbl6.2 to take the duplicates as well

# are any of these already in the synonyms?
insp_ref$synonyms[which(!is.na(match(insp_ref$synonyms$synonym.1, 
                                     nm_gn_mbl6.2))),]

```

```{r add_bl6_sce_names, message = FALSE}

# these are the IDs of gene names that are only in SCEobject
nm_gn_mbl6_ids <- rowData(mbl6_old)$ID[
  which(is.na(match(rowData(mbl6_old)$Symbol, 
                    insp_ref$gene_names$mbl6_external_gene_name)))]
length(nm_gn_mbl6_ids)

table(is.na(match(insp_ref$ids$ensembl_mbl6_id, nm_gn_mbl6_ids)))
table(is.na(match(nm_gn_mbl6_ids, insp_ref$ids$ensembl_mbl6_id)))
# all of these around 4000 IDs are in insp_ref now that I added all
# add these gene names as synonyms 
# some of these IDs are duplicated in insp_ref

insp_ref$ids[which(!is.na(match(insp_ref$ids$ensembl_mbl6_id,
                                nm_gn_mbl6_ids))),]
insp_ref$gene_names[which(!is.na(match(insp_ref$ids$ensembl_mbl6_id, 
                                       nm_gn_mbl6_ids))),]
# just add these as synonyms regardless of existence of a gene name in insp_ref

# position in insp_ref synonyms to be filled
tb_replaced_pos <- which(!is.na(match(insp_ref$ids$ensembl_mbl6_id,
                                      nm_gn_mbl6_ids)))

# positions in rowData(sce) used to fill in
rep_pos <- match(insp_ref$ids$ensembl_mbl6_id[tb_replaced_pos], 
                 rowData(mbl6_old)$ID)

insp_ref$synonyms$synonym.1[
  tb_replaced_pos] <- rowData(mbl6_old)$Symbol[rep_pos]
insp_ref$metadata$notes[tb_replaced_pos] <- paste(
  insp_ref$metadata$notes[tb_replaced_pos], 
  "Synonym_from_BL6_SCE", sep = " ")

tail(insp_ref$metadata)
table(is.na(match(rowData(mbl6_old)$Symbol, 
                  insp_ref$gene_names$mbl6_external_gene_name)))
table(is.na(match(rowData(mbl6_old)$Symbol, 
                  insp_ref$synonyms$synonym.1)))
# all symbols in these object are either in gene_names or synonyms

table(is.na(match(rowData(mbl6_bm2_unm)$Symbol, 
                  insp_ref$gene_names$mbl6_external_gene_name)))
table(is.na(match(rowData(mbl6_bm2_unm)$Symbol, 
                  insp_ref$synonyms$synonym.1)))

```

## Caroli

The Caroli objects were determined to contain identical IDs and Symbols.

### IDs

```{r compare_caroli_sce, message = FALSE}

table(is.na(match(rowData(mcar_old)$ID, 
                  insp_ref$ids$ensembl_mcar_id)))
table(is.na(match(rowData(mcar_old)$Symbol, 
                  insp_ref$gene_names$mbl6_external_gene_name)))
# good overlap of IDs but gene names overlap is quite bad

```

```{r compare_caroli_sce_ids, message = FALSE}

# there are around 300 IDs in the SCE object which are not in the insp_ref
nm_ids_car <- rowData(mcar_old)$ID[which(is.na(match(
  rowData(mcar_old)$ID, insp_ref$ids$ensembl_mcar_id)))]
length(nm_ids_car)

head(rowData(mcar_old)[which(!is.na(match(rowData(mcar_old)$ID, nm_ids_car))),])

head(nm_ids_car)
# some of these genes seem to overlap with some genes for which CAST_sc_eobject
# also unique IDs (from looking at MGI website)

```


```{r add_caroli_sce_ids, message = FALSE}

# simply add these IDs using rbind

nr_mcar <- length(nm_ids_car)
pos_mcar <- which(!is.na(match(rowData(mcar_old)$ID, nm_ids_car)))

# qa lot of the symbols to be added are actually IDs which must be removed later
length(grep("MGP", rowData(mcar_old)$Symbol))

add_mcar_sce <- list(
  ids = data.frame(
    ensembl_mbl6_id = vector("character", length = nr_mcar),
    ensembl_mcar_id = nm_ids_car,
    ensembl_mcas_id = vector("character", length = nr_mcar),
    ensembl_mspr_id = vector("character", length = nr_mcar)
  ),
  gene_names = data.frame(
    ensembl_mbl6_id = vector("character", length = nr_mcar),
    mbl6_external_gene_name = rowData(mcar_old)$Symbol[pos_mcar],
    mcar_mmusculus_homolog_associated_gene_name = rowData(
      mcar_old)$Symbol[pos_mcar],
    mcas_mmusculus_homolog_associated_gene_name = vector("character", 
                                                         length = nr_mcar),
    mspr_mmusculus_homolog_associated_gene_name = vector("character", 
                                                         length = nr_mcar)),
  synonyms = data.frame(
    ensembl_mbl6_id = vector("character", length = nr_mcar),
    mbl6_external_gene_name = rowData(mcar_old)$Symbol[pos_mcar],
    synonym.1 = vector("character", length = nr_mcar),
    synonym.2 = vector("character", length = nr_mcar),
    synonym.3 = vector("character", length = nr_mcar),
    synonym.4 = vector("character", length = nr_mcar),
    synonym.5 = vector("character", length = nr_mcar),
    synonym.6 = vector("character", length = nr_mcar)
  ),
  metadata = data.frame(
    ensembl_mbl6_id = vector("character", length = nr_mcar),
    duplicated = vector("character", length = nr_mcar),
    source = rep("CARO_SCE_object", length = nr_mcar),
    notes = vector("character", length = nr_mcar)
  )
)

nrow(add_mcar_sce$ids)

# since most of the "gene names" are actually just the IDs,
# remove all IDs 

add_mcar_sce$gene_names$mbl6_external_gene_name[grep(
  "MGP_CAROLI", 
  add_mcar_sce$gene_names$mbl6_external_gene_name)] <- paste(" ", sep = " ")
add_mcar_sce$gene_names$mcar_mmusculus_homolog_associated_gene_name[grep(
  "MGP_CAROLI", 
  add_mcar_sce$gene_names$mcar_mmusculus_homolog_associated_gene_name)] <- paste(
    " ", sep = " ")
add_mcar_sce$synonyms$mbl6_external_gene_name[grep(
  "MGP_CAROLI", 
  add_mcar_sce$synonyms$mbl6_external_gene_name)] <- paste(" ", sep = " ")

unique(add_mcar_sce$gene_names$mbl6_external_gene_name)
# only very few gene names will be added like this

```

```{r rbind_caroli_sce_ids, message = FALSE}

nrow(insp_ref$ids)
insp_ref$ids <- rbind(insp_ref$ids, add_mcar_sce$ids)
insp_ref$gene_names  <- rbind(insp_ref$gene_names, add_mcar_sce$gene_names)
insp_ref$synonyms <- rbind(insp_ref$synonyms, add_mcar_sce$synonyms)
insp_ref$metadata <- rbind(insp_ref$metadata, add_mcar_sce$metadata)
nrow(insp_ref$ids)

```

### Gene Names

```{r rcompare_caroli_sce_names, message = FALSE}

# these gene names are only in BL6_SCEobject and not in insp_ref:
nm_gn_car <- rowData(mcar_old)$Symbol[
  which(is.na(match(rowData(mcar_old)$Symbol, 
                    insp_ref$gene_names$mbl6_external_gene_name)))]
length(nm_gn_car)

# filter out instances were symbol is just the ID
nm_gn_car <- nm_gn_car[-grep("MGP_CAROLI", nm_gn_car )]
length(nm_gn_car)

# 424 gene names are missing from insp_ref

# are any of these already in the synonyms?
table(!is.na(match(nm_gn_car, insp_ref$synonyms$synonym.1)))
# yes, they are all already in synonyms
# this means they were added with data from BL6_SCE-object
# no further action required

head(insp_ref$metadata[which(!is.na(match(
  insp_ref$synonyms$synonym.1, nm_gn_car))),])
nrow(insp_ref$metadata)
```

## Castaneus

```{r compare_cast_sce, message = FALSE}

# get all info from merged and unmerged (since there were some NAs)

# ids
table(is.na(match(rowData(mcas_old)$ID, rowData(mcas_bm2_unm)$ID)))
table(is.na(match(rowData(mcas_bm2_unm)$ID, rowData(mcas_old)$ID)))

table(is.na(match(rowData(mcas_old)$ID, rowData(mcas_renorm)$ID)))
table(is.na(match(rowData(mcas_renorm)$ID, rowData(mcas_old)$ID)))

table(is.na(match(rowData(mcas_renorm)$ID, rowData(mcas_bm2_unm)$ID)))
table(is.na(match(rowData(mcas_bm2_unm)$ID, rowData(mcas_renorm)$ID)))
# mcas_old and _renorm appear identical
# mcas_old has 14 IDs mcas_bm2_unm doesn't have
# mcas_bm2_unm has 37 IDs mcas_old doesn't have

# names
table(is.na(match(rownames(mcas_old), rownames(mcas_bm2_unm))))
table(is.na(match(rownames(mcas_bm2_unm), rownames(mcas_old))))

table(is.na(match(rownames(mcas_old), rownames(mcas_renorm))))
table(is.na(match(rownames(mcas_renorm), rownames(mcas_old))))

table(is.na(match(rownames(mcas_renorm), rownames(mcas_bm2_unm))))
table(is.na(match(rownames(mcas_bm2_unm), rownames(mcas_renorm))))
# mcas_old and _renorm appear identical
# mcas_old has 3 names mcas_bm2_unm doesn't have
# mcas_bm2_unm has 26 names mcas_old doesn't have


rowData(mcas_old)$ID[which(is.na(match(rowData(mcas_old)$ID,
                                       rowData(mcas_bm2_unm)$ID)))]
rowData(mcas_bm2_unm)$ID[which(is.na(match(rowData(mcas_bm2_unm)$ID,
                                           rowData(mcas_old)$ID)))]
# mcas_bm2_unm contains the ENSMUS IDs that were added for mito genes
# use mcas_old

```

This additional genes in mcas_old are likely explaioned by a change in the 
cast reference file for cellranger.

# IDs 

```{r compare_cast_ids, message = FALSE}

# look at IDs
table(is.na(match(rowData(mcas_old)$ID, 
                  insp_ref$ids$ensembl_mcas_id)))
table(is.na(match(rowData(mcas_old)$Symbol, 
                  insp_ref$gene_names$mbl6_external_gene_name)))
# the overlap is good for IDs, not so great for gene names

length(grep("MGP", rowData(mcas_old)$Symbol))
length(grep("ENSMUS", rowData(mcas_old)$Symbol))

# remove all MGP IDs from mcas_old
rowData(mcas_old)$Symbol[grep("MGP", rowData(mcas_old)$Symbol)] <- paste("")
length(grep("MGP", rowData(mcas_old)$Symbol))

```

```{r add_cast_ids, message = FALSE}

# look the the non-matching IDs
# there are IDs in the SCE object which are not in the insp_ref
nm_ids_cast <- rowData(mcas_old)$ID[which(is.na(match(
  rowData(mcas_old)$ID, insp_ref$ids$ensembl_mcas_id)))]
length(nm_ids_cast)

nm_ids_cast2 <- rowData(mcas_bm2_unm)$ID[which(is.na(match(
  rowData(mcas_bm2_unm)$ID, insp_ref$ids$ensembl_mcas_id)))]
length(nm_ids_cast2)

# just to check again that nm_ids_cast2 only contains additional ensmus IDs
nm_ids_cast2[which(is.na(match(nm_ids_cast2, nm_ids_cast)))]

# (at least some of) these are the current IDs for the respective gene names 
# according to MGI website

```

```{r add_cast_ids2, message = FALSE}

nr_mcas <- length(nm_ids_cast)
pos_mcas <- which(!is.na(match(rowData(mcas_old)$ID,  nm_ids_cast)))

add_mcas_sce <- list(
  ids = data.frame(
    ensembl_mbl6_id = vector("character", length = nr_mcas),
    ensembl_mcar_id = vector("character", length = nr_mcas),
    ensembl_mcas_id = nm_ids_cast,
    ensembl_mspr_id = vector("character", length = nr_mcas)
  ),
  gene_names = data.frame(
    ensembl_mbl6_id = vector("character", length = nr_mcas),
    mbl6_external_gene_name = rowData(mcas_old)$Symbol[pos_mcas],
    mcar_mmusculus_homolog_associated_gene_name = vector("character",
                                                         length = nr_mcas),
    mcas_mmusculus_homolog_associated_gene_name = rowData(mcas_old)$Symbol[
      pos_mcas],
    mspr_mmusculus_homolog_associated_gene_name = vector("character", 
                                                         length = nr_mcas)
  ),
  synonyms = data.frame(
    ensembl_mbl6_id = vector("character", length = nr_mcas),
    mbl6_external_gene_name = rowData(mcas_old)$Symbol[pos_mcas],
    synonym.1 = vector("character", length = nr_mcas),
    synonym.2 = vector("character", length = nr_mcas),
    synonym.3 = vector("character", length = nr_mcas),
    synonym.4 = vector("character", length = nr_mcas),
    synonym.5 = vector("character", length = nr_mcas),
    synonym.6 = vector("character", length = nr_mcas)
   ),
  metadata = data.frame(
    ensembl_mbl6_id = vector("character", length = nr_mcas),
    duplicated = vector("character", length = nr_mcas),
    source = rep("CAST_SCEobject", length = nr_mcas),
    notes = vector("character", length = nr_mcas)
  )
)

nrow(add_mcas_sce$ids)
unique(add_mcas_sce$gene_names$mbl6_external_gene_name)
# very few gene names are actually added

```

```{r rbind_cast_ids, message = FALSE}

nrow(insp_ref$id)
insp_ref$ids <- rbind(insp_ref$ids, add_mcas_sce$ids)
insp_ref$gene_names  <- rbind(insp_ref$gene_names, add_mcas_sce$gene_names)
insp_ref$synonyms <- rbind(insp_ref$synonyms, add_mcas_sce$synonyms)
insp_ref$metadata <- rbind(insp_ref$metadata, add_mcas_sce$metadata)
tail(insp_ref$metadata)
nrow(insp_ref$id)

```

# Gene Names

```{r compare_mcas_gn, message = FALSE}

# these gene names are only in CAST_SCE_object and not in insp_ref:
nm_gn_cas1 <- rowData(mcas_old)$Symbol[which(is.na(match(
  rowData(mcas_old)$Symbol,
  insp_ref$gene_names$mbl6_external_gene_name)))]
length(nm_gn_cas1)

# check second mcas object
nm_gn_cas2 <- rowData(mcas_bm2_unm)$Symbol[which(is.na(match(
  rowData(mcas_bm2_unm)$Symbol,
  insp_ref$gene_names$mbl6_external_gene_name)))]
# filter out IDs from other SCE object as well
nm_gn_cas2 <- nm_gn_cas2[-grep("MGP", nm_gn_cas2)]
length(nm_gn_cas2)

table(is.na(match(nm_gn_cas2, nm_gn_cas1)))
table(is.na(match(nm_gn_cas1, nm_gn_cas2)))
# the additional names match, use nm_gn_cas1 because sce_old has more IDs

table(is.na(match(nm_gn_cas1[nm_gn_cas1 != ""], insp_ref$synonyms$synonym.1)))
# around 2500 of those are already in synonyms
# which means they were added from BL6_SCEobject

head(insp_ref$metadata[which(!is.na(match(insp_ref$synonyms$synonym.1,
                                     nm_gn_cas1))),])
head(insp_ref$synonym[which(!is.na(match(insp_ref$synonyms$synonym.1, 
                                    nm_gn_cas1))),])

```

```{r add_mcas_gn, message = FALSE}

# for the remaining around 550 gene names neither in gene names nor synonyms,
# add them as synonyms to the corresponding IDs

nm_gn_cas1 <- nm_gn_cas1[which(is.na(match(nm_gn_cas1, 
                                           insp_ref$synonyms$synonym.1)))] 
length(nm_gn_cas1)

# these are the IDs of gene names that are only in SCEobject
nm_gn_ids_cas <- rowData(mcas_old)$ID[which(
  !is.na(match(rowData(mcas_old)$Symbol, nm_gn_cas1)))]
length(nm_gn_ids_cas)

table(is.na(match(insp_ref$ids$ensembl_mcas_id, nm_gn_ids_cas)))
table(is.na(match(nm_gn_ids_cas, insp_ref$ids$ensembl_mcas_id)))
# all of the IDs are now in insp_ref

head(insp_ref$ids[which(!is.na(match(insp_ref$ids$ensembl_mcas_id,
                                nm_gn_ids_cas))),])
head(insp_ref$gene_names[which(!is.na(match(insp_ref$ids$ensembl_mcas_id, 
                                       nm_gn_ids_cas))),])
head(insp_ref$metadata[which(!is.na(match(insp_ref$ids$ensembl_mcas_id,
                                     nm_gn_ids_cas))),])
head(insp_ref$synonyms[which(!is.na(match(insp_ref$ids$ensembl_mcas_id, 
                                     nm_gn_ids_cas))),])
# most of those IDs already have gene names 
# (but not in the synonyms DF for some reason)
# just add these as synonyms regardless of existence of a gene name in insp_ref

```

```{r add_mcas_gn, message = FALSE}

# position in insp_ref synonyms to be filled
tb_replaced_pos <- which(!is.na(match(insp_ref$ids$ensembl_mcas_id,
                                      nm_gn_ids_cas)))
length(tb_replaced_pos)

# positions in rowData(sce) used to fill in
# match it insp_ref IDs in the required positions because some are duplicated
replace_pos <- match(insp_ref$ids$ensembl_mcas_id[tb_replaced_pos], 
                     rowData(mcas_old)$ID)
length(replace_pos)

insp_ref$synonyms$synonym.2[
  tb_replaced_pos] <- rowData(mcas_old)$Symbol[replace_pos]
insp_ref$metadata$notes[
  tb_replaced_pos] <- paste(insp_ref$metadata$notes[tb_replaced_pos], 
                            "Synonym2_from_CAST_SCE", sep = " ")

head(insp_ref$synonyms[tb_replaced_pos,])
head(insp_ref$gene_names[tb_replaced_pos,])
head(insp_ref$metadata[tb_replaced_pos,])


table(is.na(match(rowData(mcas_old)$Symbol[rowData(mcas_old)$Symbol != ""], 
                  insp_ref$gene_names$mbl6_external_gene_name)))
# after removing IDs and empty entries from symbols there are still around
# 3000 gene names left that are not found in insp_ref

table(is.na(match(rowData(mcas_old)$Symbol[rowData(mcas_old)$Symbol != ""], 
                  insp_ref$synonyms$synonym.1)))
# they are now found in synonsmy

```

## Spretus

Spretus is a bit different because of the Symbols2 col in rowData.
It was previously determined that spretus objects have identical IDs

```{r check_spret_sce1, message = FALSE}

table(is.na(match(rowData(mspr_old)$ID, insp_ref$ids$ensembl_mspr_id)))
table(is.na(match(rowData(mspr_old)$Symbol, insp_ref$ids$ensembl_mspr_id)))
table(is.na(match(rowData(mspr_old)$Symbols2, insp_ref$ids$ensembl_mspr_id)))
# in IDs/Symbol of the object, there are 393 more IDs than in insp_ref
# in Symbols 2, there are 25000 more IDs

table(is.na(match(rowData(mspr_old)$ID, 
                  insp_ref$gene_names$mbl6_external_gene_name)))
table(is.na(match(rowData(mspr_old)$Symbol, 
                  insp_ref$gene_names$mbl6_external_gene_name)))
table(is.na(match(rowData(mspr_old)$Symbols2, 
                  insp_ref$gene_names$mbl6_external_gene_name)))
# of those, 24000 are already in gene names

```

To deal with this easier, the object is changed so that IDs ONLY contains ALL 
IDs, and Symbols2 ONLY contains ALL Symbold

```{r check_spret_sce2, message = FALSE}

# Change the mspr_old object and for later comparison to the original
# use mspr_renorm as they were originally identical

table(is.na(match(rowData(mspr_old)$ID,
                  rowData(mspr_old)$Symbols2[
                    grep("MGP",rowData(mspr_old)$Symbols2)])))

table(is.na(match(rowData(mspr_old)$Symbols2[
  grep("MGP",rowData(mspr_old)$Symbols2)], rowData(mspr_old)$ID)))

# There seem to be no additional IDs in Symbols2, only gene names
# replace IDs in Symbols with ""
rowData(mspr_old)$Symbols2[grep("MGP",rowData(mspr_old)$Symbols2)] <- paste("")

# are there any non-ids in the ID row?
rowData(mspr_old)$ID[-grep("MGP", rowData(mspr_old)$ID )]

```

```{r compare_spret_sce, message = FALSE}

# IDS
table(is.na(match(rowData(mspr_old)$ID, 
                  insp_ref$ids$ensembl_mspr_id)))
# around 400 extra IDs in msor_old

table(is.na(match(rowData(mspr_old)$Symbols2, 
                  insp_ref$gene_names$mbl6_external_gene_name)))
# around 1700 extra names in mspr_old

```

### IDs

```{r compare_spret_sce_ids, message = FALSE}

# look the the non-matching IDs
nm_ids_mspr <- rowData(mspr_old)$ID[which(is.na(match(
  rowData(mspr_old)$ID, insp_ref$ids$ensembl_mspr_id)))]
length(nm_ids_mspr)

head(rowData(mspr_old)[which(!is.na(match(rowData(mspr_old)$ID, 
                                          nm_ids_mspr))),])
nrow(rowData(mspr_old))
unique(rowData(mspr_old)$Symbols2[which(!is.na(match(rowData(mspr_old)$ID, 
                                                     nm_ids_mspr)))])
# only around 20 gene names will be added like this

```

```{r add_spret_sce_ids, message = FALSE}

nr_mspr <- length(nm_ids_mspr)
pos_mspr <- which(!is.na(match(rowData(mspr_old)$ID, nm_ids_mspr)))
length(pos_mspr)

# simply add these IDs using rbind
add_mspr_sce <- list(
  ids = data.frame(
    ensembl_mbl6_id = vector("character", length = nr_mspr),
    ensembl_mcar_id = vector("character", length = nr_mspr),
    ensembl_mcas_id = vector("character", length = nr_mspr),
     ensembl_mspr_id = nm_ids_mspr
    ),
  gene_names = data.frame(
    ensembl_mbl6_id = rowData(mspr_old)$Symbols2[pos_mspr],
    mbl6_external_gene_name = vector("character", length = length(nm_ids_mspr)),
    mcar_mmusculus_homolog_associated_gene_name = vector("character", 
                                                         length = nr_mspr),
    mcas_mmusculus_homolog_associated_gene_name = vector("character", 
                                                         length = nr_mspr),
    mspr_mmusculus_homolog_associated_gene_name = rowData(
      mspr_old)$Symbols2[pos_mspr]
    ),
  synonyms = data.frame(
    ensembl_mbl6_id = vector("character", length = nr_mspr),
    mbl6_external_gene_name = rowData(mspr_old)$Symbols2[pos_mspr],
    synonym.1 = vector("character", length = nr_mspr),
    synonym.2 = vector("character", length = nr_mspr),
    synonym.3 = vector("character", length = nr_mspr),
    synonym.4 = vector("character", length = nr_mspr),
    synonym.5 = vector("character", length = nr_mspr),
    synonym.6 = vector("character", length = nr_mspr)
    ),
  metadata = data.frame(
    ensembl_mbl6_id = vector("character", length = nr_mspr),
    duplicated = vector("character", length = nr_mspr),
    source = rep("SPRET_SCEobject", length = nr_mspr),
    notes = vector("character", length = nr_mspr)
    )
)

nrow(add_mspr_sce)
head(add_mspr_sce$ids)
head(add_mspr_sce$gene_names)
head(add_mspr_sce$metadata)

```

```{r rbind_spret_sce_ids}

nrow(insp_ref$ids)
insp_ref$ids <- rbind(insp_ref$ids, add_mspr_sce$ids)
insp_ref$gene_names  <- rbind(insp_ref$gene_names, add_mspr_sce$gene_names)
insp_ref$synonyms <- rbind(insp_ref$synonyms, add_mspr_sce$synonyms)
insp_ref$metadata <- rbind(insp_ref$metadata, add_mspr_sce$metadata)
tail(insp_ref$metadata)
nrow(insp_ref$ids)

```

### Gene Names

```{r compare_spret_sce_gn, message = FALSE}

# look the the non-matching IDs
nm_gn_mspr <- rowData(mspr_old)$Symbols2[which(is.na(match(
  rowData(mspr_old)$Symbols2, insp_ref$gene_names$mbl6_external_gene_name)))]
length(nm_gn_mspr)
# around 1700 gene names are not found in insp_ref

table(is.na(match(nm_gn_mspr, insp_ref$synonyms$synonym.1)))
# apparently, all but six of those gene names are already in synonyms

# add the six remaining gene names
nm_gn_mspr <- nm_gn_mspr[which(is.na(match(nm_gn_mspr, 
                                           insp_ref$synonyms$synonym.1)))]

# get the corresponding IDs
nm_gn_mspr_ids <- rowData(mspr_old)$ID[which(!is.na(
  match(rowData(mspr_old)$Symbols2, nm_gn_mspr)))]

# check what mspr_old looks like
rowData(mspr_old)[which(!is.na(match(rowData(mspr_old)$Symbols2, nm_gn_mspr))),]

tb_replaced_mspr <- which(!is.na(match(insp_ref$ids$ensembl_mspr_id, 
                                       nm_gn_mspr_ids)))
to_replace_mspr <-  which(!is.na(match(rowData(mspr_old)$ID,
                                       nm_gn_mspr_ids)))

fill_insp <- match(insp_ref$ids$ensembl_mspr_id[tb_replaced_mspr], 
                   rowData(mspr_old)$ID[to_replace_mspr])


insp_ref$ids[which(!is.na(match(insp_ref$ids$ensembl_mspr_id,
                                               nm_gn_mspr_ids))),]
insp_ref$synonyms[which(!is.na(match(insp_ref$ids$ensembl_mspr_id,
                                               nm_gn_mspr_ids))),]

insp_ref$synonyms$synonym.3[tb_replaced_mspr][fill_insp] <- rowData(
    mspr_old)$Symbols2[to_replace_mspr]
insp_ref$metadata$notes[tb_replaced_mspr][fill_insp] <- paste(
  insp_ref$metadata$notes[tb_replaced_mspr][fill_insp], 
  "Synonym3_from_MSPR_SCE", sep = ",")
  
insp_ref$ids[tb_replaced_mspr,]
insp_ref$gene_names[tb_replaced_mspr,]
insp_ref$synonyms[tb_replaced_mspr,]
insp_ref$metadata[tb_replaced_mspr,]

```

## Checkpoint

```{r save_cp_2, message = FALSE}

list.save(insp_ref, file = "../../data/databases/insp_ref_004.rds")

```


```{r load_cp_2, message = FALSE}

insp_ref <- list.load(file = "../../data/databases/insp_ref_004.rds")

```

# Check with Perrine's sources

```{r check_gtf_ids, message = FALSE}

# are there any IDs in the gtf files that are not in insp_ref?
table(is.na(match(mcar_gtf$gene_id, insp_ref$ids$ensembl_mcar_id[
  !is.na(insp_ref$ids$ensembl_mcar_id)])))
table(is.na(match(mcas_gtf$gene_id, insp_ref$ids$ensembl_mcas_id[
  !is.na(insp_ref$ids$ensembl_mcas_id)])))
table(is.na(match(mspr_gtf$gene_id, insp_ref$ids$ensembl_mspr_id[
  !is.na(insp_ref$ids$ensembl_mspr_id)])))
# no, fortunately not

```

```{r check_gtf_gns, message = FALSE}

# are there any gene names in the gtf files that ar enot in insp_ref?
table(is.na(match(mcar_gtf$gene_name[which(!is.na(mcar_gtf$gene_name))],
                  insp_ref$gene_name$mbl6_external_gene_name[which(!is.na(
                    insp_ref$gene_name$mbl6_external_gene_name != ""))])))

table(is.na(match(mcas_gtf$gene_name[which(!is.na(mcas_gtf$gene_name))],
                  insp_ref$gene_name$mbl6_external_gene_name[which(!is.na(
                    insp_ref$gene_name$mbl6_external_gene_name != ""))])))

table(is.na(match(mspr_gtf$gene_name[which(!is.na(mspr_gtf$gene_name))],
                  insp_ref$gene_name$mbl6_external_gene_name[ which(!is.na(
                    insp_ref$gene_name$mbl6_external_gene_name != ""))])))

# there are some gene names in mcas (why am I not surprised?)
# around 3000

# what about synonyms?
table(is.na(match(mcas_gtf$gene_name[which(!is.na(mcas_gtf$gene_name))],
                  insp_ref$synonyms$synonym.1[which( !is.na(
                    insp_ref$gene_name$mbl6_external_gene_name != ""))])))
table(is.na(match(mcas_gtf$gene_name[which(!is.na(mcas_gtf$gene_name))],
                  insp_ref$synonyms$synonym.2[which( !is.na(
                    insp_ref$gene_name$mbl6_external_gene_name != ""))])))
table(is.na(match(mcas_gtf$gene_name[which(!is.na(mcas_gtf$gene_name))],
                  insp_ref$synonyms$synonym.3[which( !is.na(
                    insp_ref$gene_name$mbl6_external_gene_name != ""))])))
# around 3000 are actually in synonyms

# Add the additional gene names in synonyms 4 in case something changes later
# and denote the sources of the synonyms in the notes section
# also, some of these gene_names are NOT found in synonyms.1

# check mcas

add_gn <- mcas_gtf$gene_name[is.na(match(
  mcas_gtf$gene_name, insp_ref$gene_names$mbl6_external_gene_name[
    insp_ref$gene_names$mbl6_external_gene_name != ""]))]
table(is.na(match(add_gn, insp_ref$gene_names$mbl6_external_gene_name)))

add_gn <- add_gn[is.na(match(add_gn, insp_ref$synonyms$synonym.1))]
length(add_gn)
add_gn <- add_gn[is.na(match(add_gn, insp_ref$synonyms$synonym.2))]
length(add_gn)

# nothing to add here

```


## Checkpoint

```{r save_cp_2, message = FALSE}

list.save(insp_ref, file = "../../data/databases/insp_ref_005.rds")

```


```{r load_cp_2, message = FALSE}

insp_ref <- list.load(file = "../../data/databases/insp_ref_005.rds")

```


# Finalize

## Turn "" into NAs

```{r fill_""_nas, message = FALSE}

# For continuity 
insp_ref$ids$ensembl_mbl6_id[insp_ref$ids$ensembl_mbl6_id == ""] <- NA
insp_ref$ids$ensembl_mcar_id[insp_ref$ids$ensembl_mcar_id == ""] <- NA
insp_ref$ids$ensembl_mcas_id[insp_ref$ids$ensembl_mcas_id == ""] <- NA
insp_ref$ids$ensembl_mspr_id[insp_ref$ids$ensembl_mspr_id == ""] <- NA


insp_ref$gene_names$ensembl_mbl6_id[
  insp_ref$gene_names$ensembl_mbl6_id == ""] <- NA
insp_ref$gene_names$mbl6_external_gene_name[
  insp_ref$gene_names$mbl6_external_gene_name == ""] <- NA
insp_ref$gene_names$mcar_mmusculus_homolog_associated_gene_name[
  insp_ref$gene_names$mcar_mmusculus_homolog_associated_gene_name == ""] <- NA
insp_ref$gene_names$mcas_mmusculus_homolog_associated_gene_name[
  insp_ref$gene_names$mcas_mmusculus_homolog_associated_gene_name == ""] <- NA
insp_ref$gene_names$mspr_mmusculus_homolog_associated_gene_name[
  insp_ref$gene_names$mspr_mmusculus_homolog_associated_gene_name == ""] <- NA

insp_ref$synonyms$ensembl_mbl6_id[insp_ref$synonyms$ensembl_mbl6_id == ""] <- NA
insp_ref$metadata$ensembl_mbl6_id[insp_ref$metadata$ensembl_mbl6_id == ""] <- NA

```

## Add info on duplicates

```{r ensembl_id_duplicates, message = FALSE}

# I was really expecting to have no duplicates
table(duplicated(insp_ref$ids$ensembl_mbl6_id[which(!is.na(
  insp_ref$ids$ensembl_mbl6_id))]))
# but there are some

# which IDs are duplicated
dup_ids <- insp_ref$ids$ensembl_mbl6_id[
  which(!is.na(insp_ref$ids$ensembl_mbl6_id))][
    duplicated(insp_ref$ids$ensembl_mbl6_id[which(!is.na(
      insp_ref$ids$ensembl_mbl6_id))]) == TRUE]

# look at the rows containing duplicated IDs
insp_ref$ids[which(!is.na(match(insp_ref$ids$ensembl_mbl6_id, dup_ids))),]
# they are all CAST derived

# this is problematic because some ensembl IDs have multiple mcas IDs
# which in return also have multipe ensembl IDs
# so at the end cast IDs or ensembl ids are basically describing THE SAME THING?

# I will leave these ensembl IDs in there for reference but I really think 
# this is problematic
# so I will flag them
insp_ref$metadata$duplicated[which(!is.na(match(
  insp_ref$ids$ensembl_mbl6_id, dup_ids)))] <- paste(
    insp_ref$metadata$duplicated[which(!is.na(match(
      insp_ref$ids$ensembl_mbl6_id, dup_ids)))], "ENSEMBL_ID", sep = " ")

insp_ref$ids[which(!is.na(match(insp_ref$ids$ensembl_mbl6_id, dup_ids))),]
insp_ref$gene_names[which(!is.na(match(insp_ref$ids$ensembl_mbl6_id,dup_ids))),]
insp_ref$synonyms[which(!is.na(match(insp_ref$ids$ensembl_mbl6_id, dup_ids))),]
insp_ref$metadata[which(!is.na(match(insp_ref$ids$ensembl_mbl6_id, dup_ids))),]

```


```{r exspecies_id_duplicates, message = FALSE}

table(duplicated(insp_ref$ids$ensembl_mcar_id[
  which(!is.na(insp_ref$ids$ensembl_mcar_id))]))
table(duplicated(insp_ref$ids$ensembl_mcas_id[
  which(!is.na(insp_ref$ids$ensembl_mcas_id))]))
table(duplicated(insp_ref$ids$ensembl_mspr_id[
  which(!is.na(insp_ref$ids$ensembl_mspr_id))]))

# mcar
dup_ids_mcar <- insp_ref$ids$ensembl_mcar_id[which(!is.na(
  insp_ref$ids$ensembl_mcar_id))][duplicated(
    insp_ref$ids$ensembl_mcar_id[which(
      !is.na(insp_ref$ids$ensembl_mcar_id))]) == TRUE]

insp_ref$ids[which(!is.na(match(
  insp_ref$ids$ensembl_mcar_id, dup_ids_mcar))),]
table(table(insp_ref$ids$ensembl_mcar_id[which(!is.na(match(
  insp_ref$ids$ensembl_mcar_id, dup_ids_mcar)))]))

insp_ref$metadata$duplicated[which(!is.na(match(
  insp_ref$ids$ensembl_mcar_id, dup_ids_mcar)))] <- paste(
    insp_ref$metadata$duplicated[which(!is.na(match(
      insp_ref$ids$ensembl_mcar_id, dup_ids_mcar)))], "MCAR_ID", sep = " ")
insp_ref$metadata[which(!is.na(match(insp_ref$ids$ensembl_mcar_id, 
                                     dup_ids_mcar))),]

# mcas
dup_ids_mcas <- insp_ref$ids$ensembl_mcas_id[which(!is.na(
  insp_ref$ids$ensembl_mcas_id))][duplicated(insp_ref$ids$ensembl_mcas_id[
    which(!is.na(insp_ref$ids$ensembl_mcas_id))]) == TRUE]

insp_ref$ids[which(!is.na(match(insp_ref$ids$ensembl_mcas_id, dup_ids_mcas))),]
table(table(insp_ref$ids$ensembl_mcas_id[which(
  !is.na(match(insp_ref$ids$ensembl_mcas_id, dup_ids_mcas)))]))

insp_ref$metadata$duplicated[which(!is.na(match(
  insp_ref$ids$ensembl_mcas_id, dup_ids_mcas)))] <- paste(
    insp_ref$metadata$duplicated[which(!is.na(match(
      insp_ref$ids$ensembl_mcas_id, dup_ids_mcas)))], "MCAS_ID", sep = " ")
insp_ref$metadata[which(!is.na(match(insp_ref$ids$ensembl_mcas_id, 
                                     dup_ids_mcas))),]

# mspr
dup_ids_mspr <- insp_ref$ids$ensembl_mspr_id[which(!is.na(
  insp_ref$ids$ensembl_mspr_id))][duplicated(insp_ref$ids$ensembl_mspr_id[
    which(!is.na(insp_ref$ids$ensembl_mspr_id))]) == TRUE]

insp_ref$ids[which(!is.na(match(insp_ref$ids$ensembl_mspr_id, dup_ids_mspr))),]
table(table(insp_ref$ids$ensembl_mspr_id[which(!is.na(match(
  insp_ref$ids$ensembl_mspr_id, dup_ids_mspr)))]))

insp_ref$metadata$duplicated[which(!is.na(match(
  insp_ref$ids$ensembl_mspr_id, dup_ids_mspr)))] <- paste(
    insp_ref$metadata$duplicated[which(!is.na(match(
      insp_ref$ids$ensembl_mspr_id, dup_ids_mspr)))], "MSPR_ID", sep = " ")
insp_ref$metadata[which(!is.na(match(insp_ref$ids$ensembl_mspr_id, 
                                     dup_ids_mspr))),]

```

```{r exspecies_id_duplicates_check, message = FALSE}

pos_mcar_dup <- grep("MCAR_ID", insp_ref$metadata$duplicated)
pos_mcas_dup <- grep("MCAS_ID", insp_ref$metadata$duplicated)
pos_mspr_dup <- grep("MSPR_ID", insp_ref$metadata$duplicated)
pos_bl6_dup <- grep("ENSEMBL_ID", insp_ref$metadata$duplicated)

# duplicated IDs between exotic species overlap
table(is.na(match(pos_mcar_dup, pos_mcas_dup)))
table(is.na(match(pos_mcar_dup, pos_mspr_dup)))
table(is.na(match(pos_mcas_dup, pos_mspr_dup)))

table(is.na(match(pos_bl6_dup, pos_mcas_dup)))
# as suspected, duplicated MCAS_IDs and ENSEMBL_IDs do co-exist
insp_ref$metadata[pos_mcas_dup[which(!is.na(match(pos_mcas_dup,
                                                  pos_bl6_dup)))],]

# check if there are some corresponding MCAS_IDs in the mouse-derived part 
# of insp_ref

# these are the MCAS_IDs of rows that contain duplicated ENSEMBL IDs
mcas_ids_coinc <- insp_ref$ids$ensembl_mcas_id[pos_bl6_dup]

# print all rows containing these MCAS_ÌDS
insp_ref$ids[which(!is.na(match(insp_ref$ids$ensembl_mcas_id,
                                mcas_ids_coinc))),]
# luckily, no MCAS_IDs in the mouse-derived part of insp_ref
# is matched to an ENSMUS ID which is duplicated

# how many rows are affected by duplicates?

length(insp_ref$metadata$duplicated[insp_ref$metadata$duplicated != ""])

```

In total, 1379 rows are affected by duplicated IDs, this doesn't seem too bad 
for an object with over 90,000 rows.

Note: this is not the sum of the number of duplicated IDs per species for 
two reasons:

1) The function duplicate() only prints the duplicates, not originals, 
so the number of affected rows is higher than the printed number, since the
rows containing the originals are technically also affected.
To account for this I paid attention to use 
object[which(!is.na(match(object, duplicates)))] to also capture the originals.

2) Some duplicates affect the same row, 
decreasing the total number of rows affected.

```{r gene_name_duplicates, message = FALSE}

# I don't care about the mcar/mcas/mspr gene_names at the moment.
# I just added those for reference, not for actual use.

table(duplicated(insp_ref$gene_names$mbl6_external_gene_name[which(
  !is.na(insp_ref$gene_names$mbl6_external_gene_name))]))

dup_gn <- insp_ref$gene_names$mbl6_external_gene_name[which(!is.na(
  insp_ref$gene_names$mbl6_external_gene_name))][duplicated(
    insp_ref$gene_names$mbl6_external_gene_name[which(!is.na(
      insp_ref$gene_names$mbl6_external_gene_name))]) == TRUE]

insp_ref$gene_names[which(!is.na(match(
  insp_ref$gene_names$mbl6_external_gene_name, dup_gn))),]
table(table(insp_ref$gene_names$mbl6_external_gene_name[which(
  !is.na(match(insp_ref$gene_names$mbl6_external_gene_name, dup_gn)))]))


insp_ref$metadata$duplicated[which(!is.na(match(
  insp_ref$gene_names$mbl6_external_gene_name, dup_gn)))] <- paste(
    insp_ref$metadata$duplicated[which(!is.na(match(
      insp_ref$gene_names$mbl6_external_gene_name, dup_gn)))],
    "GENE_NAME", sep = " ")
insp_ref$metadata[which(!is.na(match(
  insp_ref$gene_names$mbl6_external_gene_name, dup_gn))),]

length(insp_ref$metadata$duplicated[grep("GENE_NAME", 
                                         insp_ref$metadata$duplicate)])

```

# Save Expanded

```{r save_prelim, message = FALSE}

list.save(insp_ref, file = "../../data/databases/insp_ref_comp.rds")

```


```{r load_prelim, message = FALSE}

insp_ref <- list.load(file = "../../data/databases/insp_ref_comp.rds")

```

